---
title: "PopHIVE Cyclosporiasis  Resources"
subtitle: "PopHIVE is a community-driven platform that empowers you with multiple sources of health data - so you can explore, understand, and act on what matters most."
date: today
from: markdown-smart
format:
  html:
    embed-resources: true
    title-block-banner: "#00356b"
    title-block-banner-color: white
    toc: true
    toc-title: "On this page"
    toc-expand: true
    code-fold: true
    fig-align: center
    smart: false
execute:
  echo: false
  warning: false
  message: false
---

```{=html}
<link rel="preconnect" href="https://fonts.googleapis.com">
<link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
<link href="https://fonts.googleapis.com/css2?family=Rubik:wght@400;500;600&family=Source+Serif+4:opsz,wght@8..60,400;8..60,600&display=swap" rel="stylesheet">
<style>
body { font-family: 'Rubik', sans-serif; color: #26323d; background-color: #ffffff; padding-top: 2.75rem; }
h1, h2, h3, h4, .title, .subtitle { font-family: 'Source Serif 4', serif; color: #00356b; font-weight: 600; }
h2 { border-bottom: 2px solid #bedeff; padding-bottom: 0.3rem; margin-top: 2.5rem; }
a { color: #00356b; }
a:hover { color: #4e98e4; }
.quarto-title-banner .title, .quarto-title-banner .subtitle, .quarto-title-banner .quarto-title-meta { color: #ffffff !important; }
.quarto-title-banner { background-color: #00356b; }
.figure-caption, figcaption { font-family: 'Rubik', sans-serif; color: #767676; font-size: 0.85rem; }
::selection { background: #bedeff; }
.source-note { font-size: 0.85rem; color: #767676; margin-top: -0.5rem; margin-bottom: 1.5rem; }
.pophive-nav {
  position: fixed; top: 0; left: 0; right: 0; z-index: 1000;
  background: #00356b; padding: 0.7rem 1.5rem; display: flex; align-items: center;
  justify-content: space-between; font-family: 'Rubik', sans-serif;
}
.pophive-nav a { color: #ffffff !important; text-decoration: none; font-size: 0.95rem; font-weight: 500; }
.pophive-nav a:hover { color: #bedeff !important; }
.pophive-brand {
  font-family: 'Source Serif 4', serif !important; font-weight: 600 !important;
  font-size: 1.15rem !important; letter-spacing: 0.02em;
}
.pophive-nav-links { display: flex; gap: 1.75rem; }
.pophive-footer {
  /* Breaks out of quarto's centered, max-width content column so the navy
     band spans the full browser width, edge to edge, like the top nav. */
  width: 100vw; position: relative; left: 50%; right: 50%;
  margin-left: -50vw; margin-right: -50vw; margin-top: 3rem;
  padding: 1.75rem 1.5rem; text-align: center;
  background: #00356b; color: #dbe7f5; font-family: 'Rubik', sans-serif;
  font-size: 0.85rem; line-height: 1.6;
}
.pophive-footer strong { font-family: 'Source Serif 4', serif; color: #ffffff; font-size: 1rem; }
.pophive-footer a { color: #bedeff !important; }
.pophive-footer a:hover { color: #ffffff !important; }
.pophive-footer .disclaimer { color: #a9bdd4; font-size: 0.8rem; margin-top: 0.4rem; }
.chart-info { margin: 0.5rem 0 2rem 0; }
.chart-info summary {
  display: inline-block; cursor: pointer; list-style: none;
  font-family: 'Rubik', sans-serif; font-size: 0.85rem; font-weight: 500;
  color: #00356b; background: #ffffff; border: 1px solid #00356b;
  border-radius: 999px; padding: 0.35rem 0.9rem; user-select: none;
}
.chart-info summary::-webkit-details-marker { display: none; }
.chart-info summary::before { content: "\24D8  "; }
.chart-info summary:hover { background: #eaf2fc; }
.chart-info[open] summary { background: #00356b; color: #ffffff; }
.chart-info p {
  margin: 0.75rem 0 0 0; font-size: 0.85rem; color: #444; max-width: 42rem;
  background: #f7f9fc; border-left: 3px solid #bedeff; padding: 0.6rem 0.9rem;
}
/* Quarto's TOC nav links default to a single line and truncate; make nested
   sub-section entries wrap onto multiple lines like normal body text. */
#TOC a { white-space: normal; word-wrap: break-word; }
</style>
```

```{=html}
<nav class="pophive-nav">
  <a href="https://www.pophive.org/" class="pophive-brand">PopHIVE</a>
  <div class="pophive-nav-links">
    <a href="https://www.pophive.org/">Home</a>
    <a href="https://www.pophive.org/about">About</a>
  </div>
</nav>
```

```{r}
#| label: setup
#| include: false

library(dplyr)
library(sf)
library(plotly)

YEAR <- 2026
MAP_YEARS <- 2022:YEAR
SCRAPER_STATE_FIPS <- c("12", "26", "39", "54", "18") # FL, MI, OH, WV, IN
state_names <- c(FL = "Florida", MI = "Michigan", OH = "Ohio", WV = "West Virginia", IN = "Indiana")

# States with zero cyclosporiasis cases reported to NNDSS in every year
# 2022-2026; shown as missing (not colored) on the national map rather than
# as a true zero, since a flat zero more likely reflects no reporting than
# no cases.
NO_REPORT_STATES <- c("Idaho", "Minnesota", "Mississippi", "Pennsylvania")

CACHE_DIR <- file.path(tempdir(), "cyclo_map_cache")
dir.create(CACHE_DIR, showWarnings = FALSE, recursive = TRUE)
as_of <- format(Sys.Date(), "%b %d, %Y")

# -----------------------------------------------------------------------------
# PopHIVE brand styling - colors and fonts matching
# https://github.com/PopHIVE/PopHivePlots (navy #00356b, Source Serif 4 /
# Rubik, viridis "magma" reversed for choropleths). All figures below are
# plotly widgets, so fonts are picked up straight from the page's Google
# Fonts <link> (no showtext/sysfonts needed - that's only for static ggplot
# output).
# -----------------------------------------------------------------------------
pophive_navy <- "#00356b"
pophive_line_colors <- c("#000004", "#29125a", "#6b1d80", "#ad347c", "#e95661", "#fe9e6d")
pophive_line_types  <- c("solid", "dashed", "solid", "dashed", "solid", "dashed")

# Shared font specs for the interactive (plotly) figures below.
plotly_title_font   <- list(family = "Source Serif 4, serif", color = pophive_navy, size = 20)
plotly_body_font    <- list(family = "Rubik, sans-serif", color = pophive_navy, size = 11)
plotly_caption_font <- list(family = "Rubik, sans-serif", color = "gray50", size = 10)

# viridis "magma" (reversed) colorscale, as a list of [fraction, hex] pairs
# for plotly's `colorscale` argument - matches the ggplot maps' brand palette.
pophive_magma_colorscale <- local({
  hex <- viridisLite::magma(32, direction = -1)
  Map(function(p, col) list(p, col), seq(0, 1, length.out = length(hex)), hex)
})
```

```{r}
#| label: fetch-data

# cyclo_resources is a private repo - clone/pull via git (uses cached GitHub
# credentials); raw.githubusercontent.com download would 404 without a token.
cyclo_dir <- file.path(CACHE_DIR, "cyclo_resources")
if (!dir.exists(cyclo_dir)) {
  system2("git", c("clone", "--depth", "1",
                    "https://github.com/PopHIVE/cyclo-resources.git", shQuote(cyclo_dir)),
          stdout = FALSE, stderr = FALSE)
} else {
  system2("git", c("-C", shQuote(cyclo_dir), "pull", "--ff-only"),
          stdout = FALSE, stderr = FALSE)
}

download_if_missing <- function(url, dest) {
  if (!file.exists(dest)) download.file(url, dest, mode = "wb", quiet = TRUE)
}

# NNDSS and Epic Health Alerts are live surveillance feeds that update weekly,
# so (unlike the static reference files below) they must be re-fetched on
# every render rather than reused once cached - otherwise CACHE_DIR silently
# pins the page to whatever snapshot happened to exist on disk.
download.file(
  "https://raw.githubusercontent.com/PopHIVE/Ingest/main/data/nnds/standard/data.csv.gz",
  file.path(CACHE_DIR, "nnds_data.csv.gz"), mode = "wb", quiet = TRUE
)
# Epic Research Health Alerts now live in the Ingest repo
# (data/epic_health_alerts, scraped there from epicresearch.org); read the
# standardized feed from Ingest like the other Ingest sources above.
download.file(
  "https://raw.githubusercontent.com/PopHIVE/Ingest/main/data/epic_health_alerts/standard/data.csv.gz",
  file.path(CACHE_DIR, "epic_health_alerts.csv.gz"), mode = "wb", quiet = TRUE
)
download_if_missing(
  "https://raw.githubusercontent.com/PopHIVE/Ingest/main/resources/all_fips.csv.gz",
  file.path(CACHE_DIR, "all_fips.csv.gz")
)
download_if_missing(
  "https://raw.githubusercontent.com/PopHIVE/Ingest/main/resources/pop_state.csv.gz",
  file.path(CACHE_DIR, "pop_state.csv.gz")
)
download_if_missing(
  "https://raw.githubusercontent.com/PopHIVE/Ingest/main/resources/census_population_2021.csv.xz",
  file.path(CACHE_DIR, "census_population_2021.csv.xz")
)
```

```{r}
#| label: prep-data
#| results: hide

# FIPS + population lookups (state, and county for FL/MI/OH/WV/IN)
all_fips <- vroom::vroom(file.path(CACHE_DIR, "all_fips.csv.gz"), show_col_types = FALSE)
state_fips <- all_fips %>%
  filter(nchar(geography) == 2) %>%
  select(geography, geography_name, state)

pop_state_raw <- vroom::vroom(file.path(CACHE_DIR, "pop_state.csv.gz"), show_col_types = FALSE) %>%
  filter(age_level == "Total") %>%
  select(geography_name = geography, popsize)

# All 50 states + DC, including FL/MI/OH/WV/IN
all_state_pop <- state_fips %>%
  inner_join(pop_state_raw, by = "geography_name") %>%
  filter(geography_name != "United States")

pop_county <- vroom::vroom(file.path(CACHE_DIR, "census_population_2021.csv.xz"), show_col_types = FALSE) %>%
  filter(nchar(GEOID) == 5, substr(GEOID, 1, 2) %in% SCRAPER_STATE_FIPS) %>%
  select(geography = GEOID, popsize = Total)

# NNDS: state-level cumulative-YTD cyclosporiasis. cyclosporiasis is ALREADY
# cumulative YTD per the source's measure_info.json, so take the latest
# reported week rather than summing.
nnds <- vroom::vroom(file.path(CACHE_DIR, "nnds_data.csv.gz"), show_col_types = FALSE)

# One cumulative-YTD snapshot per state per year, back to MAP_YEARS[1] - the
# national map's year dropdown lets users step through these.
state_data_by_year <- bind_rows(lapply(MAP_YEARS, function(yr) {
  ytd <- nnds %>%
    filter(mmwr_year == yr, !is.na(geography)) %>%
    group_by(geography) %>%
    filter(mmwr_week == max(mmwr_week)) %>%
    ungroup() %>%
    select(geography, cases = cyclosporiasis)
  all_state_pop %>%
    left_join(ytd, by = "geography") %>%
    mutate(cases = ifelse(is.na(cases), 0, cases),
           # Force perennial non-reporters to NA (missing) instead of 0, in
           # every year, so they render as unfilled on the map rather than
           # implying a confirmed zero-case count.
           cases = ifelse(geography_name %in% NO_REPORT_STATES, NA_real_, cases),
           cases_per_100k = cases / popsize * 100000,
           mmwr_year = yr)
}))

# County-level feeds for FL / OH (weekly counts -> sum for YTD total) and
# MI / WV / IN (already cumulative -> take latest snapshot per county). Left-
# join from the full county population list so non-reporting counties show as 0.
fl <- vroom::vroom(file.path(cyclo_dir, "data/fl_cyclo/standard/data.csv.gz"), show_col_types = FALSE)
fl_county <- fl %>%
  mutate(geography = sprintf("%05d", geography)) %>%
  filter(format(time, "%Y") == as.character(YEAR)) %>%
  group_by(geography) %>%
  summarize(cases = sum(fl_cyclo_cases, na.rm = TRUE) + sum(fl_cyclo_cases_probable, na.rm = TRUE), .groups = "drop")

oh <- vroom::vroom(file.path(cyclo_dir, "data/oh_cyclo/standard/data.csv.gz"), show_col_types = FALSE)
oh_county <- oh %>%
  mutate(geography = sprintf("%05d", geography)) %>%
  filter(format(time, "%Y") == as.character(YEAR)) %>%
  group_by(geography) %>%
  summarize(cases = sum(oh_cyclo_cases, na.rm = TRUE), .groups = "drop")

# Note: MI's cumulative count covers the Southeast Michigan outbreak period
# (since 2026-06-22 per MDHHS), not full-calendar-year Jan-Jun baseline cases.
mi <- vroom::vroom(file.path(cyclo_dir, "data/mi_cyclo/standard/data.csv.gz"), show_col_types = FALSE)
mi_county <- mi %>%
  mutate(geography = sprintf("%05d", geography)) %>%
  filter(format(time, "%Y") == as.character(YEAR), time == max(time)) %>%
  group_by(geography) %>%
  summarize(cases = sum(mi_cyclo_cases_cumulative, na.rm = TRUE), .groups = "drop")

# MDHHS's actual posted "Last Update" date - distinct from mi$time above,
# which is that date rolled forward to the Saturday ending its reporting week
# (see mi_cyclo's ingest.R). Read from the raw scrape snapshots, which record
# the true posted date, so this stays current automatically as MDHHS updates
# the page instead of being hardcoded here.
mi_last_update <- vroom::vroom(
  file.path(cyclo_dir, "data/mi_cyclo/raw/mi_cyclo_county_snapshots.csv"),
  show_col_types = FALSE
) %>%
  pull(last_update_date) %>%
  as.Date() %>%
  max() %>%
  format("%b %d, %Y")

# WV, like MI, only ever exposes a single cumulative-since-outbreak-start
# snapshot (OCR'd from a dashboard image; see wv_cyclo's ingest.R), not a
# full-year time series - take the latest snapshot per county, and read the
# true OEPS-posted "Last updated" date from the raw scrape snapshots (WV's
# `time` column is already that exact date, unlike MI's Saturday rollover,
# but reading it the same way keeps both blocks symmetric).
wv <- vroom::vroom(file.path(cyclo_dir, "data/wv_cyclo/standard/data.csv.gz"), show_col_types = FALSE)
wv_county <- wv %>%
  mutate(geography = sprintf("%05d", geography)) %>%
  filter(format(time, "%Y") == as.character(YEAR), time == max(time)) %>%
  group_by(geography) %>%
  summarize(cases = sum(wv_cyclo_cases_cumulative, na.rm = TRUE), .groups = "drop")

wv_last_update <- vroom::vroom(
  file.path(cyclo_dir, "data/wv_cyclo/raw/wv_cyclo_county_snapshots.csv"),
  show_col_types = FALSE
) %>%
  pull(last_update_date) %>%
  as.Date() %>%
  max() %>%
  format("%b %d, %Y")

# IN, like MI and WV, only ever exposes a single cumulative-since-
# investigation-start snapshot (a server-rendered HTML table; see in_cyclo's
# ingest.R) - take the latest snapshot per county. Unlike MI's Saturday
# rollover, IN's own `time` is already the true IDOH-posted date (IDOH updates
# daily, Mon-Fri), so in_last_update is read the same way purely to stay
# symmetric with the mi_/wv_last_update blocks above.
in_ <- vroom::vroom(file.path(cyclo_dir, "data/in_cyclo/standard/data.csv.gz"), show_col_types = FALSE)
in_county <- in_ %>%
  mutate(geography = sprintf("%05d", geography)) %>%
  filter(format(time, "%Y") == as.character(YEAR), time == max(time)) %>%
  group_by(geography) %>%
  summarize(cases = sum(in_cyclo_cases_cumulative, na.rm = TRUE), .groups = "drop")

in_last_update <- vroom::vroom(
  file.path(cyclo_dir, "data/in_cyclo/raw/in_cyclo_county_snapshots.csv"),
  show_col_types = FALSE
) %>%
  pull(last_update_date) %>%
  as.Date() %>%
  max() %>%
  format("%b %d, %Y")

county_cases <- bind_rows(fl_county, oh_county, mi_county, wv_county, in_county)

# Latest reported date within YEAR for each state's county feed - these can
# differ from each other (and from `as_of`, which is just today's date).
state_asof <- c(
  FL = format(max(fl$time[format(fl$time, "%Y") == as.character(YEAR)]), "%b %d, %Y"),
  MI = format(max(mi$time[format(mi$time, "%Y") == as.character(YEAR)]), "%b %d, %Y"),
  OH = format(max(oh$time[format(oh$time, "%Y") == as.character(YEAR)]), "%b %d, %Y"),
  WV = format(max(wv$time[format(wv$time, "%Y") == as.character(YEAR)]), "%b %d, %Y"),
  IN = format(max(in_$time[format(in_$time, "%Y") == as.character(YEAR)]), "%b %d, %Y")
)

county_data <- pop_county %>%
  left_join(county_cases, by = "geography") %>%
  mutate(cases = ifelse(is.na(cases), 0, cases),
         cases_per_100k = cases / popsize * 100000)

# County shapefile for the FL/MI/OH/WV/IN panel map below (still a static ggplot
# map, so it needs sf geometry; the national map is an interactive plotly
# choropleth and only needs state abbreviations, not shapefiles).
options(tigris_use_cache = TRUE)
Sys.setenv(TIGRIS_CACHE_DIR = file.path(CACHE_DIR, "tigris_cache"))
dir.create(Sys.getenv("TIGRIS_CACHE_DIR"), showWarnings = FALSE, recursive = TRUE)

# WGS84 (not shift_geometry(), which reprojects to a non-geographic CRS for
# repositioning AK/HI/PR - unneeded here, and plotly needs plain lon/lat).
us_counties <- tigris::counties(state = names(state_names), cb = TRUE, resolution = "20m", year = 2022) %>%
  sf::st_transform(4326) %>%
  select(GEOID, NAMELSAD, STUSPS)

county_map <- us_counties %>% left_join(county_data, by = c("GEOID" = "geography"))
```

```{r}
#| label: epic-prep
#| results: hide

# -----------------------------------------------------------------------------
# Epic Cosmos diarrhea data, pulled fresh at render time straight from the
# canonical PopHIVE/Ingest `epic_diarrhea` data source (which itself pulls
# from PopHIVE/epic_preprocessing).
# All Epic values whose suppression flag == 1 are blanked to NA (per project
# convention: suppressed counts are imputed as 5 and must not be shown).
# -----------------------------------------------------------------------------
epic_base <- "https://raw.githubusercontent.com/PopHIVE/Ingest/main/data/epic_diarrhea/standard"

# geography -> population, including national ("00") and DC. state_fips already
# carries the national row ("00" / United States), so no separate bind needed.
geo_pop <- state_fips %>%
  inner_join(pop_state_raw, by = "geography_name") %>%
  select(geography, geography_name, state, popsize)

# geography -> display name, used for the state dropdowns below.
geo_name <- setNames(geo_pop$geography_name, geo_pop$geography)

# weekly.csv.gz now carries BOTH an ED-based and an all-encounters-based
# diarrhea percentage, each by age. Blank each where its own counts were
# suppressed. (`epic_n_ed_encounters_weekly` is the total ED encounter count,
# formerly `epic_n_all_encounters_weekly`; used for the coverage share below.)
epic_weekly <- vroom::vroom(file.path(epic_base, "data_weekly.csv.gz"), show_col_types = FALSE) %>%
  mutate(
    pct_ed_diarrhea = ifelse(
      epic_suppressed_flag_ed_diarrhea == 1 | epic_suppressed_flag_ed_encounters_weekly == 1,
      NA_real_, epic_pct_ed_diarrhea),
    pct_all_diarrhea = ifelse(
      epic_suppressed_flag_all_diarrhea == 1 | epic_suppressed_flag_encounters_total_weekly == 1,
      NA_real_, epic_pct_all_diarrhea),
    # Raw numerator/denominator counts backing each percentage, for display in
    # chart tooltips. Masked to NA in lockstep with the percentage itself so a
    # suppressed count (imputed as 5 upstream) is never surfaced.
    n_ed_diarrhea = ifelse(is.na(pct_ed_diarrhea), NA_real_, epic_n_ed_diarrhea),
    n_ed_encounters = ifelse(is.na(pct_ed_diarrhea), NA_real_, epic_n_ed_encounters_weekly),
    n_all_diarrhea = ifelse(is.na(pct_all_diarrhea), NA_real_, epic_n_all_diarrhea),
    n_encounters_total = ifelse(is.na(pct_all_diarrhea), NA_real_, epic_n_encounters_total_weekly))

# Weekly cyclospora testing (replaces the monthly cyclospora feed). % tested is
# defined against all-cause diarrhea encounters: cyclospora tests / all
# diarrhea encounters. % positive is positive / tested (already in the file).
# Both are blanked where either underlying count was suppressed.
epic_wtests <- vroom::vroom(file.path(epic_base, "weekly_tests.csv.gz"), show_col_types = FALSE) %>%
  mutate(
    pct_tested = ifelse(
      epic_suppressed_flag_cyclospora_tested == 1 | epic_suppressed_flag_all_diarrhea == 1,
      NA_real_, 100 * epic_n_cyclospora_tested / epic_n_all_diarrhea),
    pct_positive = ifelse(
      epic_suppressed_flag_cyclospora_positive == 1 | epic_suppressed_flag_cyclospora_tested == 1,
      NA_real_, epic_pct_cyclospora_positive),
    # Masked raw counts backing each percentage, for tooltip display. The two
    # measures have different suppression conditions, so epic_n_cyclospora_tested
    # needs masking twice: once as % tested's numerator, once as % positive's
    # denominator.
    n_tested = ifelse(is.na(pct_tested), NA_real_, epic_n_cyclospora_tested),
    n_diarrhea_encounters = ifelse(is.na(pct_tested), NA_real_, epic_n_all_diarrhea),
    n_positive = ifelse(is.na(pct_positive), NA_real_, epic_n_cyclospora_positive),
    n_tests_denom = ifelse(is.na(pct_positive), NA_real_, epic_n_cyclospora_tested))

# Earliest date in the Epic data - the weekly time series below start here.
epic_start <- min(epic_weekly$time)

# MMWR week -> Saturday ending that week (same definition as the trend chart).
mmwr_week_ending <- function(year, week) {
  jan4 <- as.Date(paste0(year, "-01-04"))
  week1_start <- jan4 - as.integer(format(jan4, "%w"))
  week1_start + (week - 1) * 7 + 6
}

# NNDSS weekly INCIDENT cyclosporiasis cases per 100k. The source stores
# cyclosporiasis as cumulative year-to-date, so de-cumulate within each MMWR
# year (first reported week of a year is its own increment); floor at 0 so
# occasional downward data revisions don't render as negative incidence.
nnds_weekly_incident <- nnds %>%
  filter(mmwr_year >= 2022, !is.na(geography)) %>%
  select(geography, mmwr_year, mmwr_week, cum = cyclosporiasis) %>%
  inner_join(geo_pop %>% select(geography, geography_name, popsize), by = "geography") %>%
  arrange(geography, mmwr_year, mmwr_week) %>%
  group_by(geography, mmwr_year) %>%
  mutate(
    incident = pmax(cum - lag(cum, default = 0), 0),
    value_per100k = incident / popsize * 100000,
    date = mmwr_week_ending(mmwr_year, mmwr_week)
  ) %>%
  ungroup() %>%
  # States reporting zero cyclosporiasis cases in every year since 2022 are
  # shown as missing (blank) rather than a flat zero line, matching the
  # national map and the cumulative-YTD trend chart.
  mutate(value_per100k = ifelse(geography_name %in% NO_REPORT_STATES, NA_real_, value_per100k))
```

Cyclosporiasis is a diarrhea-causing infection from the parasite Cyclospira, spread through contaminated food and water. While cases typically spike during the summer months, the CDC's National Notifiable Diseases Surveillance System (NNDSS) is reporting almost 2,000 cumulative cases this year, making it the largest outbreak on record. In comparison, the U.S. typically saw about half as many by this time in previous years. Although reporting is mandated in 47 states, the exact total number of cyclosporiasis cases is difficult to pin down. Local health departments track and submit case counts to states, which then notify the CDC. Additionally, not everyone with
diarrhea seeks medical care, most people with diarrhea are not tested for cyclosporiasis, and diarrhea has many causes beyond cyclosporiasis, which complicates case tracking.

This dashboard tracks reported cases, combining CDC's NNDSS data and additional data at the county level from Florida,
Michigan, Ohio, West Virginia, and Indiana with clinical insights from Epic Cosmos. Data reflected below are current as of `r as_of`.

**Note**: State data are the authoritative source for case reporting. NNDSS data may lag or differ due to submission timing variations.

## Trends over time
National total shown by default; start typing a state name to switch to a specific state.

```{r}
#| label: trend-chart
#| results: asis

library(jsonlite)

nnds_national_weekly <- nnds %>%
  filter(geography == "00", mmwr_year >= 2022) %>%
  select(geography, mmwr_year, mmwr_week, cases = cyclosporiasis) %>%
  mutate(geography_name = "United States")

nnds_state_weekly <- nnds %>%
  filter(!is.na(geography), mmwr_year >= 2022) %>%
  select(geography, mmwr_year, mmwr_week, cases = cyclosporiasis) %>%
  inner_join(all_state_pop %>% select(geography, geography_name), by = "geography") %>%
  bind_rows(nnds_national_weekly) %>%
  arrange(geography_name, mmwr_year, mmwr_week) %>%
  mutate(
    week_ending_label = format(mmwr_week_ending(mmwr_year, mmwr_week), "%b %d"),
    # States reporting zero cyclosporiasis cases in every year since 2022 are
    # shown as missing (blank lines) rather than a flat zero, since a persistent
    # zero more likely reflects no reporting than truly no cases (matches the
    # national map's treatment).
    cases = ifelse(geography_name %in% NO_REPORT_STATES, NA_real_, cases)
  )

states_sorted <- sort(unique(nnds_state_weekly$geography_name))
years_sorted  <- sort(unique(nnds_state_weekly$mmwr_year))

# per-year (week -> "Mon DD") label map. It depends only on year + week, so it
# is stored once rather than repeated in every state's hover data.
trend_labels <- lapply(years_sorted, function(yr) {
  d <- nnds_state_weekly %>% filter(mmwr_year == yr) %>% distinct(mmwr_week, week_ending_label)
  setNames(as.list(d$week_ending_label), as.character(d$mmwr_week))
})
names(trend_labels) <- as.character(years_sorted)

# One compact record per state (drawn on demand in the browser) rather than
# ~260 pre-built, mostly-hidden plotly traces (n_states x n_years).
trend_series <- lapply(states_sorted, function(st) {
  per_year <- lapply(years_sorted, function(yr) {
    d <- nnds_state_weekly %>% filter(geography_name == st, mmwr_year == yr) %>% arrange(mmwr_week)
    list(x = d$mmwr_week, y = d$cases)
  })
  names(per_year) <- as.character(years_sorted)
  per_year
})
names(trend_series) <- states_sorted

trend_payload <- list(
  states = states_sorted,
  default = "United States",
  years = as.character(years_sorted),
  colors = pophive_line_colors[seq_along(years_sorted)],
  dashes = ifelse(pophive_line_types[seq_along(years_sorted)] == "dashed", "dash", "solid"),
  labels = trend_labels,
  series = trend_series
)

cat(sprintf('<script>window.TREND_DATA = %s;</script>\n',
            toJSON(trend_payload, auto_unbox = TRUE, na = "null", digits = 4)))
```

```{=html}
<div class="cyclo-map-controls">
  <label class="cyclo-age">State:
    <input list="trend-state-list" id="trend-state" class="cyclo-state-input" placeholder="Type a state…" autocomplete="off">
    <datalist id="trend-state-list"></datalist>
  </label>
</div>
<p class="cyclo-map-title" id="trend-title">Cumulative weekly cyclosporiasis cases</p>
<p class="cyclo-map-sub" id="trend-sub"></p>
<div id="trend-chart-div" style="width:100%;height:460px;"></div>

<script>
(function () {
  function ready() {
    if (!window.Plotly || !window.TREND_DATA) { return setTimeout(ready, 120); }
    var D = window.TREND_DATA;
    var input = document.getElementById('trend-state');
    var list = document.getElementById('trend-state-list');
    var chart = document.getElementById('trend-chart-div');
    var subEl = document.getElementById('trend-sub');
    var known = {};
    D.states.forEach(function (s) {
      var o = document.createElement('option'); o.value = s; list.appendChild(o);
      known[s.toLowerCase()] = s;
    });
    input.value = D.default;
    var last = D.default;

    function render(st) {
      var byYear = D.series[st] || {};
      var t = D.years.map(function (yr, i) {
        var s = byYear[yr] || { x: [], y: [] };
        var labs = D.labels[yr] || {};
        var cd = s.x.map(function (w) { return labs[String(w)] || ''; });
        return { x: s.x, y: s.y, customdata: cd, type: 'scatter', mode: 'lines+markers',
          name: yr, line: { color: D.colors[i], dash: D.dashes[i] }, marker: { color: D.colors[i], size: 5 },
          hovertemplate: st + ', week %{x} (week ending %{customdata}): %{y} cases<extra>' + yr + '</extra>' };
      });
      var layout = {
        margin: { t: 10, b: 45, l: 55, r: 20 }, paper_bgcolor: 'white', plot_bgcolor: 'white',
        font: { family: 'Rubik, sans-serif', color: '#00356b', size: 11 }, hovermode: 'closest',
        xaxis: { title: 'Calendar week', dtick: 4, gridcolor: 'white', zeroline: false },
        yaxis: { title: 'Cases (cumulative YTD)', gridcolor: '#e5e5e5', griddash: 'dash', zeroline: false, rangemode: 'tozero' },
        legend: { orientation: 'h', x: 0.01, y: 0.99, yanchor: 'top' }
      };
      subEl.textContent = st + ': cumulative NNDSS cases by calendar week, one line per year. Data from 2026 are preliminary.';
      Plotly.react(chart, t, layout, { displayModeBar: false, responsive: true });
    }
    function pick() { var c = known[input.value.trim().toLowerCase()]; if (c) { last = c; render(c); } }
    input.addEventListener('input', pick);
    input.addEventListener('focus', function () { input.select(); });
    input.addEventListener('change', function () {
      var c = known[input.value.trim().toLowerCase()];
      if (c) { last = c; render(c); } else { input.value = last; }
    });
    render(D.default);
  }
  ready();
})();
</script>
```

```{=html}
<details class="chart-info">
<summary>About this chart</summary>
<p>Chart shows cumulative cyclosporiasis cases reported to NNDSS by calendar week, one line per year, for the selected state (national total by default). 2026 data are preliminary and incomplete. States that have reported zero cyclosporiasis cases in every year from 2022 onward are considered 'missing' and not plotted. Not all states mandate cyclosporiasis reporting; states with zero cases may lack reporting requirements rather than having no cases. Territories are excluded.
Standardized data can be found on our <a href="https://github.com/PopHIVE/Ingest/tree/main/data/nnds/standard">GitHub</a>. The raw NNDSS data from CDC are found <a href="https://www.cdc.gov/nndss/what-is-case-surveillance/">here</a>.</p>
</details>
```

## Geographic distribution
Use the toggle to switch between cumulative cases of cyclosporiasis from CDC/NNDSS,  percent of emergency-department (ED) visits due to diarrhea from any cause from Epic Cosmos, and percent of all healthcare encounters due to diarrhea of any cause.

### Cases by state

```{r}
#| label: national-map
#| results: asis

library(jsonlite)

annual_brks <- c(0, 0.1, 0.25, 0.5, 1, 2, 3, 5)

# --- NNDSS: one slice per year (state abbr, z = log1p(cases per 100k), hover)
state_data_by_year <- state_data_by_year %>%
  mutate(hover_text = ifelse(
    is.na(cases),
    paste0(geography_name, ": missing (no cases reported since 2022)"),
    paste0(geography_name, ": ", scales::comma(cases),
           " case", ifelse(cases == 1, "", "s"),
           " (", sprintf("%.1f", cases_per_100k), " per 100,000)")
  ))
z_max_nnds <- log1p(max(state_data_by_year$cases_per_100k, na.rm = TRUE))

nnds_slices <- lapply(MAP_YEARS, function(yr) {
  d <- state_data_by_year %>% filter(mmwr_year == yr)
  list(locations = d$state,
       z    = ifelse(is.na(d$cases_per_100k), NA_real_, log1p(d$cases_per_100k)),
       text = d$hover_text)
})
names(nnds_slices) <- as.character(MAP_YEARS)

# --- Epic diarrhea: monthly by (state, age), aggregated from the weekly feed so
# BOTH an ED-based and an all-encounters-based percentage are available
# (monthly.csv.gz is ED-only). State-level only (drop "00"). Non-suppressed
# weekly counts are summed within each calendar month, then divided.
epic_map_src <- epic_weekly %>%
  filter(geography != "00") %>%
  mutate(month = as.Date(cut(time, "month"))) %>%
  group_by(geography, age, month) %>%
  summarize(
    ed_num  = sum(ifelse(epic_suppressed_flag_ed_diarrhea == 1 | epic_suppressed_flag_ed_encounters_weekly == 1, NA, epic_n_ed_diarrhea), na.rm = TRUE),
    ed_den  = sum(ifelse(epic_suppressed_flag_ed_diarrhea == 1 | epic_suppressed_flag_ed_encounters_weekly == 1, NA, epic_n_ed_encounters_weekly), na.rm = TRUE),
    all_num = sum(ifelse(epic_suppressed_flag_all_diarrhea == 1 | epic_suppressed_flag_encounters_total_weekly == 1, NA, epic_n_all_diarrhea), na.rm = TRUE),
    all_den = sum(ifelse(epic_suppressed_flag_all_diarrhea == 1 | epic_suppressed_flag_encounters_total_weekly == 1, NA, epic_n_encounters_total_weekly), na.rm = TRUE),
    .groups = "drop") %>%
  inner_join(geo_pop %>% select(geography, state, geography_name), by = "geography") %>%
  mutate(
    month_key = format(month, "%Y-%m-%d"),
    ed_pct  = ifelse(ed_den  > 0, 100 * ed_num  / ed_den,  NA_real_),
    all_pct = ifelse(all_den > 0, 100 * all_num / all_den, NA_real_)
  )

epic_ages <- intersect(
  c("Total", "<1 Years", "1-4 Years", "5-17 Years", "18-49 Years", "50-64 Years", "65+ Years"),
  unique(epic_map_src$age))
epic_months <- sort(unique(epic_map_src$month_key))

# One slice per (age, month) for a given measure ("ed"/"all").
build_epic_slices <- function(measure, label) {
  pct_col <- paste0(measure, "_pct")
  sl <- lapply(epic_ages, function(a) {
    per_month <- lapply(epic_months, function(m) {
      d <- epic_map_src %>% filter(age == a, month_key == m)
      z <- d[[pct_col]]
      hov <- ifelse(is.na(z),
        paste0(d$geography_name, ": suppressed (not shown)"),
        paste0(d$geography_name, ": ", sprintf("%.2f", z), "% of ", label, " due to diarrhea"))
      list(locations = d$state, z = ifelse(is.na(z), NA_real_, z), text = hov)
    })
    names(per_month) <- epic_months
    per_month
  })
  names(sl) <- epic_ages
  sl
}

map_payload <- list(
  colorscale = pophive_magma_colorscale,
  nnds = list(years = as.character(MAP_YEARS), slices = nnds_slices,
              zmax = z_max_nnds, tickvals = log1p(annual_brks), ticktext = annual_brks),
  epic = list(
    ages = epic_ages, months = epic_months,
    monthLabels = format(as.Date(epic_months), "%b %Y"), defaultAge = "Total",
    ed  = list(slices = build_epic_slices("ed",  "ED encounters"),
               zmax = max(epic_map_src$ed_pct,  na.rm = TRUE)),
    all = list(slices = build_epic_slices("all", "all encounters"),
               zmax = max(epic_map_src$all_pct, na.rm = TRUE))
  )
)

cat(sprintf('<script>window.CYCLO_MAP = %s;</script>\n',
            toJSON(map_payload, auto_unbox = TRUE, na = "null", digits = 6)))
```

```{=html}
<style>
.cyclo-map-controls { display: flex; flex-wrap: wrap; align-items: center; gap: 1.25rem; margin: 0.25rem 0 0.5rem; font-family: 'Rubik', sans-serif; }
.cyclo-toggle { display: inline-flex; border: 1px solid #00356b; border-radius: 999px; overflow: hidden; }
.cyclo-toggle button { border: 0; background: #ffffff; color: #00356b; font-family: 'Rubik', sans-serif; font-size: 0.85rem; font-weight: 500; padding: 0.4rem 1rem; cursor: pointer; }
.cyclo-toggle button.active { background: #00356b; color: #ffffff; }
.cyclo-age { font-size: 0.85rem; color: #00356b; font-weight: 500; }
.cyclo-age select { font-family: 'Rubik', sans-serif; font-size: 0.85rem; color: #26323d; border: 1px solid #bedeff; border-radius: 6px; padding: 0.25rem 0.4rem; margin-left: 0.35rem; }
.cyclo-age input.cyclo-state-input { font-family: 'Rubik', sans-serif; font-size: 0.85rem; color: #26323d; border: 1px solid #bedeff; border-radius: 6px; padding: 0.3rem 0.5rem; margin-left: 0.35rem; min-width: 13rem; }
.cyclo-age input.cyclo-state-input:focus { outline: none; border-color: #4e98e4; box-shadow: 0 0 0 2px #bedeff; }
.cyclo-map-title { font-family: 'Source Serif 4', serif; color: #00356b; font-weight: 600; font-size: 1.15rem; margin: 0.5rem 0 0; }
.cyclo-map-sub { font-family: 'Rubik', sans-serif; color: #767676; font-size: 0.82rem; margin: 0.15rem 0 0.25rem; }
.cyclo-slider-row { display: flex; align-items: center; gap: 0.75rem; margin: 0.25rem 0 0; font-family: 'Rubik', sans-serif; font-size: 0.82rem; color: #00356b; }
.cyclo-slider-row input[type=range] { flex: 1; accent-color: #00356b; }
#cyclo-period-label { min-width: 6.5rem; font-weight: 500; }
</style>

<div class="cyclo-map-wrap">
  <div class="cyclo-map-controls">
    <div class="cyclo-toggle">
      <button data-mode="nnds" class="active" type="button">NNDSS cyclosporiasis cases</button>
      <button data-mode="epic_ed" type="button">Epic Cosmos ED visits</button>
      <button data-mode="epic_all" type="button">Epic Cosmos all encounters</button>
    </div>
    <label class="cyclo-age" id="cyclo-age-wrap" style="display:none;">Age group:
      <select id="cyclo-age-select"></select>
    </label>
  </div>
  <p class="cyclo-map-title" id="cyclo-map-title"></p>
  <p class="cyclo-map-sub" id="cyclo-map-sub"></p>
  <div id="cyclo-map" style="width:100%;height:460px;"></div>
  <div class="cyclo-slider-row">
    <span id="cyclo-period-label"></span>
    <input type="range" id="cyclo-period" min="0" max="1" value="0" step="1" aria-label="time period">
  </div>
</div>

<script>
(function () {
  function ready() {
    if (!window.Plotly || !window.CYCLO_MAP) { return setTimeout(ready, 120); }
    var M = window.CYCLO_MAP;
    var mapDiv = document.getElementById('cyclo-map');
    var titleEl = document.getElementById('cyclo-map-title');
    var subEl = document.getElementById('cyclo-map-sub');
    var slider = document.getElementById('cyclo-period');
    var periodLabel = document.getElementById('cyclo-period-label');
    var ageWrap = document.getElementById('cyclo-age-wrap');
    var ageSelect = document.getElementById('cyclo-age-select');
    var buttons = document.querySelectorAll('.cyclo-toggle button');

    // populate age dropdown (Epic mode)
    M.epic.ages.forEach(function (a) {
      var o = document.createElement('option');
      o.value = a; o.textContent = a; ageSelect.appendChild(o);
    });
    ageSelect.value = M.epic.defaultAge;

    var state = { mode: 'nnds', age: M.epic.defaultAge, idx: M.nnds.years.length - 1 };

    // Drop entries whose z is null so those states render unfilled (missing /
    // suppressed) rather than being colored as a real low value.
    function clean(slice) {
      var loc = [], z = [], txt = [];
      for (var i = 0; i < slice.locations.length; i++) {
        if (slice.z[i] === null || slice.z[i] === undefined) continue;
        loc.push(slice.locations[i]); z.push(slice.z[i]); txt.push(slice.text[i]);
      }
      return { locations: loc, z: z, text: txt };
    }

    function current() {
      if (state.mode === 'nnds') {
        var yr = M.nnds.years[state.idx];
        return {
          slice: clean(M.nnds.slices[yr]),
          zmax: M.nnds.zmax,
          colorbar: { title: { text: 'Cases per<br>100,000' }, tickvals: M.nnds.tickvals, ticktext: M.nnds.ticktext, len: 1, thickness: 18 },
          title: 'Cyclosporiasis cases by state (NNDSS)',
          sub: 'Cumulative cases per 100,000 &middot; ' + yr,
          periodLabel: 'Year: ' + yr
        };
      }
      var measure = (state.mode === 'epic_all') ? 'all' : 'ed';
      var em = M.epic[measure];
      var mo = M.epic.months[state.idx];
      var isAll = measure === 'all';
      return {
        slice: clean(em.slices[state.age][mo]),
        zmax: em.zmax,
        colorbar: { title: { text: isAll ? '% of all<br>encounters' : '% of ED<br>encounters' }, len: 1, thickness: 18 },
        title: isAll ? 'Diarrhea across all encounters by state (Epic Cosmos)' : 'Diarrhea in ED encounters by state (Epic Cosmos)',
        sub: '% of ' + (isAll ? 'all healthcare encounters' : 'emergency-department encounters') + ' due to diarrhea &middot; Age ' + state.age + ' &middot; ' + M.epic.monthLabels[state.idx],
        periodLabel: M.epic.monthLabels[state.idx]
      };
    }

    function render(redraw) {
      var c = current();
      var trace = {
        type: 'choropleth', locationmode: 'USA-states',
        locations: c.slice.locations, z: c.slice.z, text: c.slice.text,
        hoverinfo: 'text', zmin: 0, zmax: c.zmax,
        colorscale: M.colorscale, showscale: true,
        marker: { line: { color: 'white', width: 0.5 } },
        colorbar: c.colorbar
      };
      var layout = {
        geo: { scope: 'usa', projection: { type: 'albers usa' }, bgcolor: 'transparent' },
        margin: { t: 6, b: 6, l: 0, r: 0 }, paper_bgcolor: 'white',
        font: { family: 'Rubik, sans-serif', color: '#00356b', size: 11 }
      };
      titleEl.textContent = c.title;
      subEl.innerHTML = c.sub;
      periodLabel.textContent = c.periodLabel;
      if (redraw) {
        Plotly.newPlot(mapDiv, [trace], layout, { displayModeBar: false, responsive: true });
      } else {
        Plotly.react(mapDiv, [trace], layout, { displayModeBar: false, responsive: true });
      }
    }

    function setSlider() {
      var n = (state.mode === 'nnds') ? M.nnds.years.length : M.epic.months.length;
      slider.max = n - 1;
      state.idx = n - 1;      // default to the most recent period
      slider.value = state.idx;
    }

    buttons.forEach(function (b) {
      b.addEventListener('click', function () {
        if (b.dataset.mode === state.mode) return;
        var wasNnds = state.mode === 'nnds';
        buttons.forEach(function (x) { x.classList.remove('active'); });
        b.classList.add('active');
        state.mode = b.dataset.mode;
        var isNnds = state.mode === 'nnds';
        ageWrap.style.display = isNnds ? 'none' : 'inline-block';
        // Only reset the time slider when the period domain changes (annual
        // NNDSS <-> monthly Epic); switching between the two Epic modes keeps
        // the selected month.
        if (wasNnds !== isNnds) setSlider();
        render(false);
      });
    });
    ageSelect.addEventListener('change', function () { state.age = ageSelect.value; render(false); });
    slider.addEventListener('input', function () { state.idx = parseInt(slider.value, 10); render(false); });

    setSlider();
    render(true);
  }
  ready();
})();
</script>
```

```{=html}
<details class="chart-info">
<summary>About this chart</summary>
<p><b>NNDSS view:</b> cumulative cyclosporiasis cases reported to NNDSS by state for the selected year (cases per 100,000). 2026 data are preliminary and incomplete. States that have reported zero cyclosporiasis cases in every year from 2022 onward are considered 'missing' and not plotted. Not all states mandate cyclosporiasis reporting; states with zero cases may lack reporting requirements rather than having no cases. Territories are excluded.
Standardized data are on our <a href="https://github.com/PopHIVE/Ingest/tree/main/data/nnds/standard">GitHub</a>; the CDC presents its NNDSS data <a href="https://www.cdc.gov/nndss/what-is-case-surveillance/">here</a>.<br><br>
<b>Epic diarrhea view:</b> percent of ED visits and all healthcare encounters attributable to all-cause diarrhea (ICD-10-CM A00-A09, R19.7). Insights are drawn from clinical electronic health records (EHR)from facilities using the Epic EHR and participating in Epic Cosmos. Epic Cosmos captures varying amounts of the population in different regions, so the data may be mroe representative of diarrhea in the community in some areas than others. This is a measure of diarrhea activity in the ED, not a cyclosporiasis case count. Counts fewer than 10 are suppressed and excluded (shown unfilled). Standardized data are on our <a href="https://github.com/PopHIVE/cyclo-resources/tree/main/data/cosmos-diarrhea">GitHub</a>.</p>
</details>
```

```{r}
#| label: national-map-state-totals
#| results: asis

fips_to_abbr <- setNames(names(state_names), SCRAPER_STATE_FIPS)

state_totals <- county_data %>%
  mutate(abbr = fips_to_abbr[substr(geography, 1, 2)]) %>%
  group_by(abbr) %>%
  summarize(cases = sum(cases), popsize = sum(popsize), .groups = "drop") %>%
  arrange(match(abbr, c("MI", "OH", "FL", "WV", "IN"))) %>%
  transmute(
    State = paste0(state_names[abbr], ifelse(abbr %in% c("MI", "WV", "IN"), "*", "")),
    `Number of cases` = scales::comma(cases),
    `Cases per 100,000` = sprintf("%.1f", cases / popsize * 100000),
    # state_asof["MI"]/state_asof["WV"]/state_asof["IN"] are rolled-forward
    # reporting-week dates (per mi_cyclo's/wv_cyclo's/in_cyclo's ingest
    # conventions), not the source health department's own posted date -
    # override with the actual posted dates (mi_/wv_/in_last_update) here.
    `Last updated` = case_when(
      abbr == "MI" ~ mi_last_update,
      abbr == "WV" ~ wv_last_update,
      abbr == "IN" ~ in_last_update,
      TRUE ~ state_asof[abbr]
    )
  )

cat("As noted above, data from NNDSS for this year may not be the most up-to-date. Below are totals reported by states for 2026:\n\n")
cat(knitr::kable(state_totals, format = "pipe"), sep = "\n")
cat("\n\n\\*Michigan's count reflects cases reported since July 1, 2026, West Virginia's count reflects cases reported since the start of its outbreak investigation, and Indiana's count reflects cases reported since the start of its investigation (May 1, 2026) - none of these three are a full-year total.\n")
```

### MI, OH, FL, WV, and IN: county-level detail

These five states display case counts at the county level collected from state health departments.
Hover over a county to see its case count. Most states only report cases at the state level.
Oregon also reports at the county level, but its data are suppressed due to small sample sizes because Oregon is not currently experiencing a large outbreak.


```{r}
#| label: county-maps-setup
#| include: false

# plotly has no built-in US-counties location mode, so build a small GeoJSON
# from the tigris county shapes (FL/MI/OH/WV/IN only) and use it directly.
counties_geojson_file <- tempfile(fileext = ".geojson")
sf::st_write(county_map %>% select(GEOID), counties_geojson_file, driver = "GeoJSON", quiet = TRUE)
counties_geojson <- jsonlite::fromJSON(counties_geojson_file, simplifyVector = FALSE)

county_hover <- county_map %>%
  sf::st_drop_geometry() %>%
  mutate(hover_text = paste0(NAMELSAD, ": ", scales::comma(cases),
                              " case", ifelse(cases == 1, "", "s"),
                              " (", sprintf("%.1f", cases_per_100k), " per 100,000)"))

brks <- c(0, 1, 5, 25, 100)
z_max_county <- log1p(max(county_hover$cases_per_100k, na.rm = TRUE))

state_sources <- c(
  FL = paste0("Source: Florida Dept. of Health."),
  MI = paste0("Source: Michigan Dept. of Health & Human Services."),
  OH = paste0("Source: Ohio Dept. of Health."),
  WV = paste0("Source: WV Dept. of Health, Office of Epidemiology and Prevention Services."),
  IN = paste0("Source: Indiana Dept. of Health, Infectious Disease Epidemiology & Prevention Division.")
)

# FL's feed is explicitly preliminary (subject to revision for 1-2 weeks per
# its README); MI/OH are not flagged that way, hence the differing wording.
# WV's county counts are OCR'd from a dashboard image (see wv_cyclo's
# README.md) and carry the same outbreak-cumulative framing as MI's. IN's
# county counts share that same outbreak-cumulative framing, but IDOH updates
# them daily (Mon-Fri) rather than weekly - the finest resolution of any
# source here.
state_subtitles <- c(
  FL = paste0("Cumulative year-to-date preliminary cyclosporiasis cases per 100,000 population as of ", state_asof[["FL"]], "."),
  MI = paste0("Cumulative preliminary cyclosporiasis cases per 100,000 population reported during the outbreak investigation (as of ", mi_last_update, ", starting July 1, 2026)."),
  OH = paste0("Cumulative year-to-date preliminary cyclosporiasis cases per 100,000 population as of ", state_asof[["OH"]], "."),
  WV = paste0("Cumulative preliminary cyclosporiasis cases per 100,000 population reported during the outbreak investigation (as of ", wv_last_update, ")."),
  IN = paste0("Cumulative preliminary cyclosporiasis cases per 100,000 population reported during the investigation (as of ", in_last_update, ", starting May 1, 2026).")
)

make_county_plot <- function(abbr) {
  d <- county_hover %>% filter(STUSPS == abbr)
  # Explicit height: inside the tagList wrapper the chunk's fig-height no longer
  # sizes the widget, so set it here to keep the maps their previous size.
  plot_ly(height = 500) %>%
    add_trace(
      type = "choropleth", geojson = counties_geojson, featureidkey = "properties.GEOID",
      locations = d$GEOID, z = log1p(d$cases_per_100k), zmin = 0, zmax = z_max_county,
      colorscale = pophive_magma_colorscale, text = d$hover_text, hoverinfo = "text",
      marker = list(line = list(color = "white", width = 0.4)),
      colorbar = list(
        title = list(text = "Cases per<br>100,000", font = plotly_body_font),
        tickvals = log1p(brks), ticktext = brks, tickfont = plotly_body_font,
        len = 0.8, thickness = 18
      )
    ) %>%
    layout(
      geo = list(scope = "usa", fitbounds = "locations", visible = FALSE),
      annotations = list(list(
        text = state_sources[[abbr]], xref = "paper", yref = "paper",
        x = 0, y = -0.06, xanchor = "left", showarrow = FALSE, font = plotly_caption_font
      )),
      margin = list(t = 10, b = 40),
      paper_bgcolor = "white"
    ) %>%
    config(displayModeBar = FALSE)
}

# Title + subtitle are rendered as HTML above the map (not in the plotly title)
# so the longer Michigan subtitle wraps naturally with tight line spacing,
# matching the other charts' subtitles, instead of being clipped.
county_block <- function(abbr) {
  header <- sprintf(
    '<p class="cyclo-map-title" style="margin-top:1rem;">%s</p><p class="cyclo-map-sub">%s</p>',
    state_names[[abbr]], state_subtitles[[abbr]]
  )
  htmltools::tagList(htmltools::HTML(header), make_county_plot(abbr))
}
```


```{r}
#| label: county-mi
#| fig-width: 9
#| fig-height: 7
county_block("MI")
```


```{r}
#| label: county-oh
#| fig-width: 9
#| fig-height: 7
county_block("OH")
```


```{r}
#| label: county-fl
#| fig-width: 9
#| fig-height: 7
county_block("FL")
```

```{r}
#| label: county-wv
#| fig-width: 9
#| fig-height: 7
county_block("WV")
```

```{r}
#| label: county-in
#| fig-width: 9
#| fig-height: 7
county_block("IN")
```

```{=html}
<details class="chart-info">
<summary>About these charts</summary>
<p>Cyclosporiasis  cases are obtained from individual state health departments. Rates are calculated using county-specific population denominators from the 2021 census. Standardized data can be found on our <a href="https://github.com/PopHIVE/cyclo-resources/tree/main/data/oh_cyclo">GitHub</a>. Raw data from FL are found <a href="https://www.flhealthcharts.gov/ChartsReports/rdPage.aspx?rdReport=FrequencyMerlin.Frequency">here</a>; raw data from MI are found <a href="https://www.michigan.gov/mdhhs/keep-mi-healthy/infectious-diseases/infectious-disease-outbreaks">here</a>; raw data from OH are found <a href="https://data.ohio.gov/wps/portal/gov/data/view/summary-of-infectious-diseases-in-ohio">here</a>; raw data from WV are found <a href="https://oeps.wv.gov/cyclosporiasis-outbreak">here</a> (county counts are OCR'd from a dashboard image, since WV publishes no structured data table); and raw data from IN are found <a href="https://www.in.gov/health/idepd/diseases-and-conditions-resource-page/cyclosporiasis/#Cases_by_County">here</a>.</p>
</details>
```


## Epic Cosmos Trends

::: {.callout-warning}
The latest available week from Epic Cosmos (shown with a grayed bar) is incompletedata based on small numbers and should be interpreted with caution.
:::

Epic Cosmos captures broader diarrhea activity in health care facilities and how often clinicians order a cyclospora lab test. These are population-level signals of disease activity, not case counts. National total is shown by default. The last week of data, denoted with a gray box, is incomplete and likely to change. Start typing a state name to switch to a specific state.

### Epic Cosmos diarrhea and Cyclospora testing trends
Start typing a state name to filter (national total by default). Hover any point for the underlying value. The last week of data, denoted with a gray box, is incomplete and likely to change.

```{r}
#| label: epic-nnds-compare
#| results: asis

library(jsonlite)

# Geographies in the Epic weekly feed (national + states). Dropdown lists them
# alphabetically but defaults to the United States (sorted first).
compare_geos <- epic_weekly %>%
  distinct(geography) %>%
  mutate(geography_name = unname(geo_name[geography])) %>%
  filter(!is.na(geography_name)) %>%
  arrange(geography_name != "United States", geography_name)

# Epic weekly diarrhea percentages (Total age): ED-based and all-encounters
# based, plus the Epic population-coverage share (ED encounters / population).
epic_wk_total <- epic_weekly %>%
  filter(age == "Total") %>%
  left_join(geo_pop %>% select(geography, popsize), by = "geography") %>%
  mutate(share_covered = 100 * epic_n_ed_encounters_weekly / popsize)

cmp_xmax <- max(c(max(epic_wk_total$time), max(nnds_weekly_incident$date)))
r4 <- function(v) round(v, 4)

# Emit one compact record per geography and let the browser draw only the
# selected geography's handful of traces on demand, rather than shipping ~160
# mostly-hidden plotly traces. Keeps full interactivity at a fraction of the
# page weight. The bottom-panel testing measures are now weekly.
cmp_series <- lapply(compare_geos$geography, function(g) {
  dn <- nnds_weekly_incident %>% filter(geography == g) %>% arrange(date)
  de <- epic_wk_total %>% filter(geography == g) %>% arrange(time)
  dt <- epic_wtests %>% filter(geography == g) %>% arrange(time)
  out <- list(
    nnds     = list(x = format(dn$date), y = r4(dn$value_per100k)),
    diar     = list(x = format(de$time), y = r4(de$pct_ed_diarrhea), share = r4(de$share_covered),
                     num = de$n_ed_diarrhea, den = de$n_ed_encounters),
    diar_all = list(x = format(de$time), y = r4(de$pct_all_diarrhea),
                     num = de$n_all_diarrhea, den = de$n_encounters_total),
    tested   = list(x = format(dt$time), y = r4(dt$pct_tested),
                     num = dt$n_tested, den = dt$n_diarrhea_encounters),
    # Positive counts are suppressed in most state-weeks; plot every
    # non-suppressed point (gaps where suppressed) rather than gating on a
    # minimum number of weeks.
    positive = list(x = format(dt$time), y = r4(dt$pct_positive),
                     num = dt$n_positive, den = dt$n_tests_denom)
  )
  out
})
names(cmp_series) <- compare_geos$geography

cmp_payload <- list(
  geos = lapply(seq_len(nrow(compare_geos)), function(i)
    list(code = compare_geos$geography[i], name = compare_geos$geography_name[i])),
  default = "00",
  xrange = c(format(epic_start), format(cmp_xmax)),
  colors = list(nnds = pophive_navy, diar = "#ad347c", diar_all = "#fe9e6d",
                tested = "#2a7f62", positive = "#6b1d80"),
  series = cmp_series
)

cat(sprintf('<script>window.CMP_DATA = %s;</script>\n',
            toJSON(cmp_payload, auto_unbox = TRUE, na = "null", digits = 4)))
```

```{=html}
<div class="cyclo-map-controls">
  <label class="cyclo-age">State:
    <input list="cmp-state-list" id="cmp-state" class="cyclo-state-input" placeholder="Type a state…" autocomplete="off">
    <datalist id="cmp-state-list"></datalist>
  </label>
</div>
<p class="cyclo-map-title">Epic diarrhea trends compared to cyclosporiasis cases (NNDSS)</p>
<p class="cyclo-map-sub">Five panels share the same time range; each has its own y-axis scale.</p>
<p class="cyclo-map-title" style="font-size:1rem;margin-top:0.75rem;">Epic Cosmos: % of ED encounters due to diarrhea</p>
<div id="cmp-chart-ed" style="width:100%;height:190px;"></div>
<p class="cyclo-map-title" style="font-size:1rem;margin-top:1rem;">Epic Cosmos: % of all encounters due to diarrhea</p>
<div id="cmp-chart-all" style="width:100%;height:190px;"></div>
<p class="cyclo-map-title" style="font-size:1rem;margin-top:1rem;">Cyclosporiasis cases (NNDSS)</p>
<div id="cmp-chart-nnds" style="width:100%;height:190px;"></div>
<p class="cyclo-map-title" style="font-size:1rem;margin-top:1rem;">Stool and cyclospora testing: % of diarrhea encounters tested for cyclospora</p>
<div id="cmp-chart-tested" style="width:100%;height:190px;"></div>
<p class="cyclo-map-title" style="font-size:1rem;margin-top:1rem;">Stool and cyclospora testing: % of cyclospora tests positive</p>
<p class="cyclo-map-sub">Positive Cyclospora test counts are suppressed in a large majority of the states and are not plotted if suppressed.</p>
<div id="cmp-chart-positive" style="width:100%;height:190px;"></div>

<script>
(function () {
  function ready() {
    if (!window.Plotly || !window.CMP_DATA) { return setTimeout(ready, 120); }
    var D = window.CMP_DATA;
    var input = document.getElementById('cmp-state');
    var list = document.getElementById('cmp-state-list');
    var edDiv = document.getElementById('cmp-chart-ed');
    var allDiv = document.getElementById('cmp-chart-all');
    var nndsDiv = document.getElementById('cmp-chart-nnds');
    var testedDiv = document.getElementById('cmp-chart-tested');
    var positiveDiv = document.getElementById('cmp-chart-positive');
    var nameByCode = {}, codeByName = {};
    D.geos.forEach(function (g) {
      var o = document.createElement('option'); o.value = g.name; list.appendChild(o);
      nameByCode[g.code] = g.name; codeByName[g.name.toLowerCase()] = g.code;
    });
    input.value = nameByCode[D.default];
    var last = D.default;

    var baseFont = { family: 'Rubik, sans-serif', color: '#00356b', size: 11 };
    var headFont = { family: 'Source Serif 4, serif', color: '#00356b', size: 13 };
    var cfg = { displayModeBar: false, responsive: true };
    // Matched l/r margins on all four panels so their plot areas line up
    // horizontally even though they are separate charts. All four panels use
    // the same D.xrange for their x-axis, so the time axes stay aligned
    // (matching x-axes) while each panel keeps its own independent y-axis.
    var LR = { l: 66, r: 66 };

    // Epic Cosmos' most recent week is still accruing data when the site
    // rebuilds, so shade the last interval to flag it as incomplete rather
    // than let it read as a real dip/rise.
    function lastWeekShapes(xArr) {
      if (!xArr || xArr.length < 2) return [];
      var lastMs = new Date(xArr[xArr.length - 1]).getTime();
      var prevMs = new Date(xArr[xArr.length - 2]).getTime();
      var pad = (lastMs - prevMs) * 0.35;
      return [{ type: 'rect', xref: 'x', yref: 'paper', y0: 0, y1: 1,
        x0: new Date(prevMs - pad).toISOString().slice(0, 10),
        x1: new Date(lastMs + pad).toISOString().slice(0, 10),
        fillcolor: 'rgba(90,90,90,0.32)', line: { width: 0 }, layer: 'below' }];
    }

    // Pairs numerator/denominator arrays into per-point [num, den] tuples for
    // use as Plotly customdata, so tooltips can show the counts behind a %.
    function zip(a, b) {
      if (!a || !b) return [];
      return a.map(function (v, i) { return [v, b[i]]; });
    }

    function render(code) {
      var s = D.series[code], nm = nameByCode[code];

      // ---- Panel 1: Epic % of ED encounters due to diarrhea ----
      var ed = [];
      if (s.diar) ed.push({ x: s.diar.x, y: s.diar.y, customdata: zip(s.diar.num, s.diar.den), type: 'scatter', mode: 'lines',
        name: '% of ED encounters due to diarrhea', line: { color: D.colors.diar, width: 2 }, connectgaps: false,
        hovertemplate: nm + '<br>week ending %{x|%b %d, %Y}<br><b>%{y:.2f}%</b> of ED encounters were diarrhea' +
          '<br>%{customdata[0]:,.0f} of %{customdata[1]:,.0f} ED encounters<extra></extra>' });
      var edLayout = {
        margin: { t: 10, b: 34, l: LR.l, r: LR.r }, paper_bgcolor: 'white', plot_bgcolor: 'white',
        font: baseFont, hovermode: 'closest', showlegend: false,
        // Every panel shows its own time axis, all sharing D.xrange (same range,
        // same plot width/margins), so tick positions line up across panels
        // even though NNDSS's data (panel 3) extends further than Epic's.
        xaxis: { type: 'date', range: D.xrange, gridcolor: '#eee', zeroline: false },
        yaxis: { title: 'Percent due to diarrhea', color: D.colors.diar, gridcolor: '#e5e5e5', griddash: 'dash', zeroline: false, rangemode: 'tozero' },
        shapes: lastWeekShapes(s.diar && s.diar.x)
      };

      // ---- Panel 2: Epic % of all encounters due to diarrhea ----
      var all = [];
      if (s.diar_all) all.push({ x: s.diar_all.x, y: s.diar_all.y, customdata: zip(s.diar_all.num, s.diar_all.den), type: 'scatter', mode: 'lines',
        name: '% of all encounters due to diarrhea', line: { color: D.colors.diar_all, width: 2 }, connectgaps: false,
        hovertemplate: nm + '<br>week ending %{x|%b %d, %Y}<br><b>%{y:.2f}%</b> of all encounters were diarrhea-related' +
          '<br>%{customdata[0]:,.0f} of %{customdata[1]:,.0f} encounters<extra></extra>' });
      var allLayout = {
        margin: { t: 10, b: 34, l: LR.l, r: LR.r }, paper_bgcolor: 'white', plot_bgcolor: 'white',
        font: baseFont, hovermode: 'closest', showlegend: false,
        xaxis: { type: 'date', range: D.xrange, gridcolor: '#eee', zeroline: false },
        yaxis: { title: 'Percent due to diarrhea', color: D.colors.diar_all, gridcolor: '#e5e5e5', griddash: 'dash', zeroline: false, rangemode: 'tozero' },
        shapes: lastWeekShapes(s.diar_all && s.diar_all.x)
      };

      // ---- Panel 3: NNDSS cyclosporiasis cases ----
      var nnds = [];
      if (s.nnds) nnds.push({ x: s.nnds.x, y: s.nnds.y, type: 'scatter', mode: 'lines',
        name: 'NNDSS cases per 100k', line: { color: D.colors.nnds, width: 2 },
        hovertemplate: nm + '<br>week ending %{x|%b %d, %Y}<br><b>%{y:.3f}</b> NNDSS incident cases per 100,000<extra></extra>' });
      var nndsLayout = {
        margin: { t: 10, b: 34, l: LR.l, r: LR.r }, paper_bgcolor: 'white', plot_bgcolor: 'white',
        font: baseFont, hovermode: 'closest', showlegend: false,
        xaxis: { type: 'date', range: D.xrange, gridcolor: '#eee', zeroline: false },
        yaxis: { title: 'NNDSS cases per 100k', color: D.colors.nnds, gridcolor: '#e5e5e5', griddash: 'dash', zeroline: false }
      };

      // ---- Panel 4a: % of diarrhea encounters tested for cyclospora ----
      var tested = [];
      if (s.tested) tested.push({ x: s.tested.x, y: s.tested.y, customdata: zip(s.tested.num, s.tested.den), type: 'scatter', mode: 'lines+markers',
        name: '% of diarrhea encounters tested for cyclospora', line: { color: D.colors.tested, width: 1.8 },
        marker: { color: D.colors.tested, size: 4 }, connectgaps: false,
        hovertemplate: nm + '<br>week ending %{x|%b %d, %Y}<br><b>%{y:.2f}%</b> of diarrhea encounters had a cyclospora test' +
          '<br>%{customdata[0]:,.0f} of %{customdata[1]:,.0f} diarrhea encounters<extra></extra>' });
      var testedLayout = {
        margin: { t: 10, b: 34, l: LR.l, r: LR.r }, paper_bgcolor: 'white', plot_bgcolor: 'white',
        font: baseFont, hovermode: 'closest', showlegend: false,
        xaxis: { type: 'date', range: D.xrange, gridcolor: '#eee', zeroline: false },
        yaxis: { title: '% tested', color: D.colors.tested, gridcolor: '#e5e5e5', griddash: 'dash', zeroline: false, rangemode: 'tozero' },
        shapes: lastWeekShapes(s.tested && s.tested.x)
      };

      // ---- Panel 4b: % of cyclospora tests positive ----
      var positive = [];
      if (s.positive) positive.push({ x: s.positive.x, y: s.positive.y, customdata: zip(s.positive.num, s.positive.den), type: 'scatter', mode: 'lines+markers',
        name: '% cyclospora positive', line: { color: D.colors.positive, width: 1.5, dash: 'dash' },
        marker: { color: D.colors.positive, size: 4 }, connectgaps: false,
        hovertemplate: nm + '<br>week ending %{x|%b %d, %Y}<br><b>%{y:.2f}%</b> of cyclospora tests positive' +
          '<br>%{customdata[0]:,.0f} of %{customdata[1]:,.0f} tests<extra></extra>' });
      var positiveLayout = {
        margin: { t: 10, b: 34, l: LR.l, r: LR.r }, paper_bgcolor: 'white', plot_bgcolor: 'white',
        font: baseFont, hovermode: 'closest', showlegend: false,
        xaxis: { type: 'date', range: D.xrange, gridcolor: '#eee', zeroline: false },
        yaxis: { title: '% positive', color: D.colors.positive, gridcolor: '#e5e5e5', griddash: 'dash', zeroline: false, rangemode: 'tozero' },
        shapes: lastWeekShapes(s.positive && s.positive.x)
      };

      Plotly.react(edDiv, ed, edLayout, cfg);
      Plotly.react(allDiv, all, allLayout, cfg);
      Plotly.react(nndsDiv, nnds, nndsLayout, cfg);
      Plotly.react(testedDiv, tested, testedLayout, cfg);
      Plotly.react(positiveDiv, positive, positiveLayout, cfg);
    }
    function pick() { var c = codeByName[input.value.trim().toLowerCase()]; if (c) { last = c; render(c); } }
    input.addEventListener('input', pick);
    input.addEventListener('focus', function () { input.select(); });
    input.addEventListener('change', function () {
      var c = codeByName[input.value.trim().toLowerCase()];
      if (c) { last = c; render(c); } else { input.value = nameByCode[last]; }
    });
    render(D.default);
  }
  ready();
})();
</script>
```

```{=html}
<details class="chart-info">
<summary>About these charts</summary>
<p>
<b>NNDSS data:</b> cumulative cyclosporiasis cases reported to NNDSS by state for the selected year (cases per 100,000). NNDSS cyclosporiasis is reported as a cumulative year-to-date count, so it is de-cumulated into weekly incident cases. 2026 data are preliminary and incomplete. Not all states mandate cyclosporiasis reporting; states with zero cases may lack reporting requirements rather than having no cases. States that have reported zero cyclosporiasis cases in every year from 2022 onward are considered 'missing' and not plotted. Territories are excluded. Standardized data are on our <a href="https://github.com/PopHIVE/Ingest/tree/main/data/nnds/standard">GitHub</a>; the CDC presents its NNDSS data <a href="https://www.cdc.gov/nndss/what-is-case-surveillance/">here</a>.<br><br>
<b>Epic insights:</b> percent of ED visits and all healthcare encounters attributable to all-cause diarrhea (ICD-10-CM A00-A09, R19.7). Insights are drawn from clinical electronic health records (EHR)from facilities using the Epic EHR and participating in Epic Cosmos. Epic Cosmos captures varying amounts of the population in different regions, so the data may be more representative of diarrhea in the community in some areas than others. This is a measure of diarrhea activity, not a cyclosporiasis case count. Counts fewer than 10 are suppressed and excluded (shown unfilled). Standardized data are on our <a href="https://github.com/PopHIVE/cyclo-resources/tree/main/data/cosmos-diarrhea">GitHub</a>; the weekly file is <a href="https://github.com/PopHIVE/cyclo-resources/blob/main/data/cosmos-diarrhea/standard/weekly.csv.gz">weekly.csv.gz</a>.<br><br>
Cyclospora testing insights are also drawn from Epic Cosmos by searching for the following tests: Cyclospora SP AG in stool by immunoassay 38403-2, acid fast stain 10659-1, safranin stain 50313-6, cayetanensis presence 10850-6, cayetanensis DNA presence by NAA with non-probe detection 82206-4, and cayetanensis DNA presence by NAA with non-probe detection 41436-7. The two weekly cyclospora testing measures have different denominators: <b>% of diarrhea encounters tested</b> for cyclospora (cyclospora tests ÷ all-cause diarrhea encounters that week) and <b>% of cyclospora tests</b> that are positive (positive results ÷ tests). All Epic values with a suppressed count (fewer than 10) are excluded. The light gray shading on the four Epic panels marks the most recent week, which is still accruing data and is incomplete.
</p>
</details>
```

### Epic diarrhea by age group
Weekly share of emergency-department and all healthcare encounters attributable to all-cause diarrhea. National total shown by default; start typing a state name to switch to a specific state. The last week of data, denoted with a gray box, is incomplete and likely to change.

```{r}
#| label: epic-age-trend
#| results: asis

age_levels <- intersect(
  c("<1 Years", "1-4 Years", "5-17 Years", "18-49 Years", "50-64 Years", "65+ Years", "Total"),
  unique(epic_weekly$age))
# magma ramp for the age bands; "Total" overridden to navy + drawn thicker so
# it reads as the reference line rather than just another band.
age_pal <- setNames(viridisLite::magma(length(age_levels) + 1, direction = -1)[-1], age_levels)
age_pal["Total"] <- pophive_navy

epic_wk_age <- epic_weekly %>%
  left_join(geo_pop %>% select(geography, popsize), by = "geography") %>%
  mutate(share_covered = 100 * epic_n_ed_encounters_weekly / popsize)

# Shared weekly x-grid: store the dates once, then per geography x age two
# aligned y arrays - the ED-based and all-encounters-based diarrhea percentages
# - plus the coverage share (NA where suppressed/missing). A browser toggle
# switches which metric is drawn.
age_x <- sort(unique(epic_wk_age$time))
age_series <- lapply(compare_geos$geography, function(g) {
  per_age <- lapply(age_levels, function(a) {
    d <- epic_wk_age %>% filter(geography == g, age == a)
    ed <- rep(NA_real_, length(age_x)); all <- rep(NA_real_, length(age_x)); sh <- rep(NA_real_, length(age_x))
    ed_num <- rep(NA_real_, length(age_x)); ed_den <- rep(NA_real_, length(age_x))
    all_num <- rep(NA_real_, length(age_x)); all_den <- rep(NA_real_, length(age_x))
    idx <- match(d$time, age_x)
    ed[idx] <- d$pct_ed_diarrhea; all[idx] <- d$pct_all_diarrhea; sh[idx] <- d$share_covered
    ed_num[idx] <- d$n_ed_diarrhea; ed_den[idx] <- d$n_ed_encounters
    all_num[idx] <- d$n_all_diarrhea; all_den[idx] <- d$n_encounters_total
    list(ed = round(ed, 4), all = round(all, 4), share = round(sh, 4),
         ed_num = ed_num, ed_den = ed_den, all_num = all_num, all_den = all_den)
  })
  names(per_age) <- age_levels
  per_age
})
names(age_series) <- compare_geos$geography

age_payload <- list(
  geos = lapply(seq_len(nrow(compare_geos)), function(i)
    list(code = compare_geos$geography[i], name = compare_geos$geography_name[i])),
  default = "00",
  defaultMetric = "ed",
  ages = age_levels,
  colors = as.list(setNames(unname(age_pal[age_levels]), age_levels)),
  widths = as.list(setNames(ifelse(age_levels == "Total", 3, 1.6), age_levels)),
  x = format(age_x),
  xrange = c(format(epic_start), format(max(age_x))),
  series = age_series
)

cat(sprintf('<script>window.AGE_DATA = %s;</script>\n',
            toJSON(age_payload, auto_unbox = TRUE, na = "null", digits = 4)))
```

```{=html}
<div class="cyclo-map-controls">
  <div class="cyclo-toggle" id="age-metric-toggle">
    <button data-metric="ed" class="active" type="button">% of ED visits</button>
    <button data-metric="all" type="button">% of all encounters</button>
  </div>
  <label class="cyclo-age">State:
    <input list="age-state-list" id="age-state" class="cyclo-state-input" placeholder="Type a state…" autocomplete="off">
    <datalist id="age-state-list"></datalist>
  </label>
</div>
<p class="cyclo-map-title" id="age-title">Epic diarrhea by age group</p>
<p class="cyclo-map-sub" id="age-sub"></p>
<div id="age-chart" style="width:100%;height:480px;"></div>

<script>
(function () {
  function ready() {
    if (!window.Plotly || !window.AGE_DATA) { return setTimeout(ready, 120); }
    var D = window.AGE_DATA;
    var input = document.getElementById('age-state');
    var list = document.getElementById('age-state-list');
    var chart = document.getElementById('age-chart');
    var subEl = document.getElementById('age-sub');
    var buttons = document.querySelectorAll('#age-metric-toggle button');
    var nameByCode = {}, codeByName = {};
    D.geos.forEach(function (g) {
      var o = document.createElement('option'); o.value = g.name; list.appendChild(o);
      nameByCode[g.code] = g.name; codeByName[g.name.toLowerCase()] = g.code;
    });
    input.value = nameByCode[D.default];
    var state = { code: D.default, metric: D.defaultMetric };

    function render() {
      var s = D.series[state.code], nm = nameByCode[state.code];
      var isEd = state.metric === 'ed';
      var pop = isEd ? 'ED encounters' : 'all encounters';
      var numKey = isEd ? 'ed_num' : 'all_num';
      var denKey = isEd ? 'ed_den' : 'all_den';
      var t = D.ages.map(function (a) {
        var cd = s[a][numKey].map(function (v, i) { return [v, s[a][denKey][i]]; });
        return { x: D.x, y: s[a][state.metric], customdata: cd, type: 'scatter', mode: 'lines',
          name: a, line: { color: D.colors[a], width: D.widths[a] }, connectgaps: false,
          hovertemplate: nm + ' — ' + a + '<br>week ending %{x|%b %d, %Y}<br><b>%{y:.2f}%</b> of ' + pop + ' were diarrhea' +
            '<br>%{customdata[0]:,.0f} of %{customdata[1]:,.0f} ' + pop + '<extra></extra>' };
      });
      var layout = {
        margin: { t: 10, b: 40, l: 55, r: 20 }, paper_bgcolor: 'white', plot_bgcolor: 'white',
        font: { family: 'Rubik, sans-serif', color: '#00356b', size: 11 }, hovermode: 'closest',
        xaxis: { type: 'date', range: D.xrange, gridcolor: '#eee', zeroline: false },
        yaxis: { title: '% of ' + pop + ' (diarrhea)', gridcolor: '#e5e5e5', griddash: 'dash', zeroline: false, rangemode: 'tozero' },
        legend: { orientation: 'h', x: 0, y: -0.16 },
        // Epic Cosmos' most recent week is still accruing data at rebuild
        // time, so shade the last interval to flag it as incomplete.
        shapes: (function () {
          if (!D.x || D.x.length < 2) return [];
          var lastMs = new Date(D.x[D.x.length - 1]).getTime();
          var prevMs = new Date(D.x[D.x.length - 2]).getTime();
          var pad = (lastMs - prevMs) * 0.35;
          return [{ type: 'rect', xref: 'x', yref: 'paper', y0: 0, y1: 1,
            x0: new Date(prevMs - pad).toISOString().slice(0, 10),
            x1: new Date(lastMs + pad).toISOString().slice(0, 10),
            fillcolor: 'rgba(90,90,90,0.32)', line: { width: 0 }, layer: 'below' }];
        })()
      };
      subEl.textContent = nm + ' — % of ' + pop + ' that were all-cause diarrhea, by age, weekly.';
      Plotly.react(chart, t, layout, { displayModeBar: false, responsive: true });
    }
    function pickState() { var c = codeByName[input.value.trim().toLowerCase()]; if (c) { state.code = c; render(); } }
    input.addEventListener('input', pickState);
    input.addEventListener('focus', function () { input.select(); });
    input.addEventListener('change', function () {
      var c = codeByName[input.value.trim().toLowerCase()];
      if (c) { state.code = c; render(); } else { input.value = nameByCode[state.code]; }
    });
    buttons.forEach(function (b) {
      b.addEventListener('click', function () {
        if (b.dataset.metric === state.metric) return;
        buttons.forEach(function (x) { x.classList.remove('active'); });
        b.classList.add('active');
        state.metric = b.dataset.metric;
        render();
      });
    });
    render();
  }
  ready();
})();
</script>
```

```{=html}
<details class="chart-info">
<summary>About this chart</summary>
<p>The percent of ED visits and all healthcare encounters attributable to all-cause diarrhea (ICD-10-CM A00-A09, R19.7) are obtained from facilities using the Epic electronic health records platform and participating in Epic Cosmos. Use the toggle to switch the denominator between ED visits and all healthcare encounters. Epic Cosmos captures varying amounts of the population in different regions, so the data may be more representative of diarrhea in the community in some areas than others. This is a measure of diarrhea activity, not a cyclosporiasis case count. "Total" (navy) is the all-ages line. Counts fewer than 10 are suppressed and excluded, which is why some state/age lines have gaps. Standardized data are on our <a href="https://github.com/PopHIVE/cyclo-resources/tree/main/data/cosmos-diarrhea">GitHub</a>; the weekly file is <a href="https://github.com/PopHIVE/cyclo-resources/blob/main/data/cosmos-diarrhea/standard/weekly.csv.gz">weekly.csv.gz</a>. The light gray shading marks the most recent week, which is still accruing data and is incomplete.</p>
</details>
```

```{r}
#| label: epic-alerts-data

# Epic Research Health Alerts now live in the Ingest repo (moved there from
# epic_preprocessing); the standardized feed is downloaded from Ingest in the
# fetch-data chunk above, alongside the NNDSS data.
epic_alerts <- vroom::vroom(
  file.path(CACHE_DIR, "epic_health_alerts.csv.gz"),
  col_types = vroom::cols(geography = vroom::col_character(), .default = vroom::col_guess())
)

# The feed carries every alerted condition (see the `condition` column); keep
# only Cyclosporiasis. It accumulates one weekly snapshot per upstream run, so
# the currently-active alerts are the rows at the latest `time`.
epic_cyclo_alerts <- epic_alerts %>% filter(condition == "Cyclosporiasis")

# State rows (2-digit FIPS) are labeled "State-Wide"; county rows (5-digit
# FIPS) are labeled with the county's full name (e.g. "Wayne County"), joined
# back to its state's full name via `state` (abbreviation) - `state_fips` and
# `all_fips` are already loaded above (prep-data chunk).
epic_state_lookup <- state_fips %>%
  transmute(geography, State = geography_name, Location = "State-Wide")

epic_county_lookup <- all_fips %>%
  filter(nchar(geography) == 5) %>%
  select(geography, Location = geography_name, state) %>%
  left_join(state_fips %>% select(state, State = geography_name), by = "state") %>%
  select(geography, State, Location)

epic_geo_lookup <- bind_rows(epic_state_lookup, epic_county_lookup)

# Source dates are MM-DD-YYYY; display onset in the page's own "M/D/YYYY" style.
fmt_epic_date <- function(d) {
  d <- as.Date(d, "%m-%d-%Y")
  paste(as.integer(format(d, "%m")), as.integer(format(d, "%d")), format(d, "%Y"), sep = "/")
}

latest_alert_time <- if (nrow(epic_cyclo_alerts) > 0) {
  max(as.Date(epic_cyclo_alerts$time, "%m-%d-%Y"))
} else {
  as.Date(NA)
}
epic_current <- epic_cyclo_alerts %>%
  filter(as.Date(time, "%m-%d-%Y") == latest_alert_time) %>%
  left_join(epic_geo_lookup, by = "geography") %>%
  transmute(
    State, Location,
    `Estimated Onset` = fmt_epic_date(estimated_onset),
    `Cases per 100k (Latest Week)` = paste0(
      sprintf("%.1f", value),
      ifelse(partial_week_flag == 1, "*", "")
    )
  ) %>%
  arrange(State, Location)

# `date_epic_updated` is the Health Alerts page's own page-wide "Last updated"
# date as of the latest scrape (MM-DD-YYYY, applies to every condition).
epic_last_updated <- {
  d <- max(as.Date(epic_alerts$date_epic_updated, "%m-%d-%Y"))
  paste0(format(d, "%B"), " ", as.integer(format(d, "%d")), ", ", format(d, "%Y"))
}

epic_states <- unique(epic_current$State)
epic_states_text <- if (length(epic_states) <= 1) {
  paste0("**", epic_states, "**")
} else {
  paste0(
    paste0("**", epic_states[-length(epic_states)], "**", collapse = ", "),
    " and **", epic_states[length(epic_states)], "**"
  )
}

epic_intro_text <- if (nrow(epic_current) > 0) {
  paste0(
    "Epic Health Alerts has identified ", epic_states_text,
    " as currently experiencing elevated rates of cyclosporiasis, with cases significantly ",
    "above historical levels for this time of year. Data are current through ", epic_last_updated, "."
  )
} else {
  paste0(
    "Epic Health Alerts is not currently flagging any locations for elevated cyclosporiasis rates. ",
    "Data are current through ", epic_last_updated, "."
  )
}
```

### Epic Health Alerts: elevated cyclosporiasis detected
`r epic_intro_text`

```{r}
#| label: epic-alerts-table
#| results: asis
if (nrow(epic_current) > 0) cat(knitr::kable(epic_current, format = "pipe"), sep = "\n")
```

##### Stay Updated
These numbers are preliminary and subject to change as more data arrive. Visit the [Epic Health Alerts dashboard](https://www.epicresearch.org/health-alerts/) regularly for the latest information.

```{=html}
<details class="chart-info">
<summary>How alerts are triggered</summary>
<p> Health Alerts are generated based on analysis in Epic Cosmos, based on data from Epic Cosmos participating organizations. They do not reflect complete case counts for a condition but rather evaluate population level trends of disease activity. The data can be viewed at https://www.epicresearch.org/health-alerts/. <br>
<p>Epic Health Alerts flag conditions when three criteria are met:<br>
1. Current rates exceed those from the same period last year<br>
2. The proportion of affected patients shows accelerating growth<br>
3. The increase is statistically significant compared to the prior three years<br>
See more in Epic Health Alerts' <a href="https://www.epicresearch.org/health-alerts/#methodology">methodology section</a>.</p>
</details>
```



```{=html}
<footer class="pophive-footer">
  <p><strong>PopHIVE</strong> - Population Health Integrated Visualization Engine</p>
  <p>Yale School of Public Health &amp; Yale DISSC - Learn more at <a href="https://www.pophive.org/">pophive.org</a></p>
  <p class="disclaimer">Research tool for public health professionals. Not intended for clinical decision-making.</p>
</footer>
```
