---
title: "PopHIVE Cyclosporiasis  Resources"
subtitle: "PopHIVE is a community-driven platform that empowers you with multiple sources of health data - so you can explore, understand, and act on what matters most."
date: today
from: markdown-smart
format:
  html:
    embed-resources: true
    title-block-banner: "#00356b"
    title-block-banner-color: white
    toc: true
    toc-title: "On this page"
    code-fold: true
    fig-align: center
    smart: false
execute:
  echo: false
  warning: false
  message: false
---

```{=html}
<link rel="preconnect" href="https://fonts.googleapis.com">
<link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
<link href="https://fonts.googleapis.com/css2?family=Rubik:wght@400;500;600&family=Source+Serif+4:opsz,wght@8..60,400;8..60,600&display=swap" rel="stylesheet">
<style>
body { font-family: 'Rubik', sans-serif; color: #26323d; background-color: #ffffff; padding-top: 2.75rem; }
h1, h2, h3, h4, .title, .subtitle { font-family: 'Source Serif 4', serif; color: #00356b; font-weight: 600; }
h2 { border-bottom: 2px solid #bedeff; padding-bottom: 0.3rem; margin-top: 2.5rem; }
a { color: #00356b; }
a:hover { color: #4e98e4; }
.quarto-title-banner .title, .quarto-title-banner .subtitle, .quarto-title-banner .quarto-title-meta { color: #ffffff !important; }
.quarto-title-banner { background-color: #00356b; }
.figure-caption, figcaption { font-family: 'Rubik', sans-serif; color: #767676; font-size: 0.85rem; }
::selection { background: #bedeff; }
.source-note { font-size: 0.85rem; color: #767676; margin-top: -0.5rem; margin-bottom: 1.5rem; }
.pophive-nav {
  position: fixed; top: 0; left: 0; right: 0; z-index: 1000;
  background: #00356b; padding: 0.7rem 1.5rem; display: flex; align-items: center;
  justify-content: space-between; font-family: 'Rubik', sans-serif;
}
.pophive-nav a { color: #ffffff !important; text-decoration: none; font-size: 0.95rem; font-weight: 500; }
.pophive-nav a:hover { color: #bedeff !important; }
.pophive-brand {
  font-family: 'Source Serif 4', serif !important; font-weight: 600 !important;
  font-size: 1.15rem !important; letter-spacing: 0.02em;
}
.pophive-nav-links { display: flex; gap: 1.75rem; }
.pophive-footer {
  /* Breaks out of quarto's centered, max-width content column so the navy
     band spans the full browser width, edge to edge, like the top nav. */
  width: 100vw; position: relative; left: 50%; right: 50%;
  margin-left: -50vw; margin-right: -50vw; margin-top: 3rem;
  padding: 1.75rem 1.5rem; text-align: center;
  background: #00356b; color: #dbe7f5; font-family: 'Rubik', sans-serif;
  font-size: 0.85rem; line-height: 1.6;
}
.pophive-footer strong { font-family: 'Source Serif 4', serif; color: #ffffff; font-size: 1rem; }
.pophive-footer a { color: #bedeff !important; }
.pophive-footer a:hover { color: #ffffff !important; }
.pophive-footer .disclaimer { color: #a9bdd4; font-size: 0.8rem; margin-top: 0.4rem; }
.chart-info { margin: 0.5rem 0 2rem 0; }
.chart-info summary {
  display: inline-block; cursor: pointer; list-style: none;
  font-family: 'Rubik', sans-serif; font-size: 0.85rem; font-weight: 500;
  color: #00356b; background: #ffffff; border: 1px solid #00356b;
  border-radius: 999px; padding: 0.35rem 0.9rem; user-select: none;
}
.chart-info summary::-webkit-details-marker { display: none; }
.chart-info summary::before { content: "\24D8  "; }
.chart-info summary:hover { background: #eaf2fc; }
.chart-info[open] summary { background: #00356b; color: #ffffff; }
.chart-info p {
  margin: 0.75rem 0 0 0; font-size: 0.85rem; color: #444; max-width: 42rem;
  background: #f7f9fc; border-left: 3px solid #bedeff; padding: 0.6rem 0.9rem;
}
/* Quarto's TOC nav links default to a single line and truncate; make nested
   sub-section entries wrap onto multiple lines like normal body text. */
#TOC a { white-space: normal; word-wrap: break-word; }
</style>
```

```{=html}
<nav class="pophive-nav">
  <a href="https://www.pophive.org/" class="pophive-brand">PopHIVE</a>
  <div class="pophive-nav-links">
    <a href="https://www.pophive.org/">Home</a>
    <a href="https://www.pophive.org/about">About</a>
  </div>
</nav>
```

```{r}
#| label: setup
#| include: false

library(dplyr)
library(sf)
library(plotly)

YEAR <- 2026
MAP_YEARS <- 2022:YEAR
SCRAPER_STATE_FIPS <- c("12", "26", "39") # FL, MI, OH
state_names <- c(FL = "Florida", MI = "Michigan", OH = "Ohio")

BASE_DIR  <- "C:/Users/as5325/Desktop"
CACHE_DIR <- file.path(BASE_DIR, "cyclo_map_cache")
dir.create(CACHE_DIR, showWarnings = FALSE, recursive = TRUE)
as_of <- format(Sys.Date(), "%b %d, %Y")

# -----------------------------------------------------------------------------
# PopHIVE brand styling - colors and fonts matching
# https://github.com/PopHIVE/PopHivePlots (navy #00356b, Source Serif 4 /
# Rubik, viridis "magma" reversed for choropleths). All figures below are
# plotly widgets, so fonts are picked up straight from the page's Google
# Fonts <link> (no showtext/sysfonts needed - that's only for static ggplot
# output).
# -----------------------------------------------------------------------------
pophive_navy <- "#00356b"
pophive_line_colors <- c("#000004", "#29125a", "#6b1d80", "#ad347c", "#e95661", "#fe9e6d")
pophive_line_types  <- c("solid", "dashed", "solid", "dashed", "solid", "dashed")

# Shared font specs for the interactive (plotly) figures below.
plotly_title_font   <- list(family = "Source Serif 4, serif", color = pophive_navy, size = 20)
plotly_body_font    <- list(family = "Rubik, sans-serif", color = pophive_navy, size = 11)
plotly_caption_font <- list(family = "Rubik, sans-serif", color = "gray50", size = 10)

# viridis "magma" (reversed) colorscale, as a list of [fraction, hex] pairs
# for plotly's `colorscale` argument - matches the ggplot maps' brand palette.
pophive_magma_colorscale <- local({
  hex <- viridisLite::magma(32, direction = -1)
  Map(function(p, col) list(p, col), seq(0, 1, length.out = length(hex)), hex)
})
```

```{r}
#| label: fetch-data
#| cache: true

# cyclo_resources is a private repo - clone/pull via git (uses cached GitHub
# credentials); raw.githubusercontent.com download would 404 without a token.
cyclo_dir <- file.path(CACHE_DIR, "cyclo_resources")
if (!dir.exists(cyclo_dir)) {
  system2("git", c("clone", "--depth", "1",
                    "https://github.com/PopHIVE/cyclo_resources.git", shQuote(cyclo_dir)),
          stdout = FALSE, stderr = FALSE)
} else {
  system2("git", c("-C", shQuote(cyclo_dir), "pull", "--ff-only"),
          stdout = FALSE, stderr = FALSE)
}

download_if_missing <- function(url, dest) {
  if (!file.exists(dest)) download.file(url, dest, mode = "wb", quiet = TRUE)
}
download_if_missing(
  "https://raw.githubusercontent.com/PopHIVE/Ingest/main/data/nnds/standard/data.csv.gz",
  file.path(CACHE_DIR, "nnds_data.csv.gz")
)
download_if_missing(
  "https://raw.githubusercontent.com/PopHIVE/Ingest/main/resources/all_fips.csv.gz",
  file.path(CACHE_DIR, "all_fips.csv.gz")
)
download_if_missing(
  "https://raw.githubusercontent.com/PopHIVE/Ingest/main/resources/pop_state.csv.gz",
  file.path(CACHE_DIR, "pop_state.csv.gz")
)
download_if_missing(
  "https://raw.githubusercontent.com/PopHIVE/Ingest/main/resources/census_population_2021.csv.xz",
  file.path(CACHE_DIR, "census_population_2021.csv.xz")
)
```

```{r}
#| label: prep-data

# FIPS + population lookups (state, and county for FL/MI/OH)
all_fips <- vroom::vroom(file.path(CACHE_DIR, "all_fips.csv.gz"), show_col_types = FALSE)
state_fips <- all_fips %>%
  filter(nchar(geography) == 2) %>%
  select(geography, geography_name, state)

pop_state_raw <- vroom::vroom(file.path(CACHE_DIR, "pop_state.csv.gz"), show_col_types = FALSE) %>%
  filter(age_level == "Total") %>%
  select(geography_name = geography, popsize)

# All 50 states + DC, including FL/MI/OH
all_state_pop <- state_fips %>%
  inner_join(pop_state_raw, by = "geography_name") %>%
  filter(geography_name != "United States")

pop_county <- vroom::vroom(file.path(CACHE_DIR, "census_population_2021.csv.xz"), show_col_types = FALSE) %>%
  filter(nchar(GEOID) == 5, substr(GEOID, 1, 2) %in% SCRAPER_STATE_FIPS) %>%
  select(geography = GEOID, popsize = Total)

# NNDS: state-level cumulative-YTD cyclosporiasis. cyclosporiasis is ALREADY
# cumulative YTD per the source's measure_info.json, so take the latest
# reported week rather than summing.
nnds <- vroom::vroom(file.path(CACHE_DIR, "nnds_data.csv.gz"), show_col_types = FALSE)

# One cumulative-YTD snapshot per state per year, back to MAP_YEARS[1] - the
# national map's year dropdown lets users step through these.
state_data_by_year <- bind_rows(lapply(MAP_YEARS, function(yr) {
  ytd <- nnds %>%
    filter(mmwr_year == yr, !is.na(geography)) %>%
    group_by(geography) %>%
    filter(mmwr_week == max(mmwr_week)) %>%
    ungroup() %>%
    select(geography, cases = cyclosporiasis)
  all_state_pop %>%
    left_join(ytd, by = "geography") %>%
    mutate(cases = ifelse(is.na(cases), 0, cases),
           cases_per_100k = cases / popsize * 100000,
           mmwr_year = yr)
}))

# County-level feeds for FL / OH (weekly counts -> sum for YTD total) and MI
# (already cumulative -> take latest snapshot per county). Left-join from the
# full county population list so non-reporting counties show as 0.
fl <- vroom::vroom(file.path(cyclo_dir, "data/fl_cyclo/standard/data.csv.gz"), show_col_types = FALSE)
fl_county <- fl %>%
  mutate(geography = sprintf("%05d", geography)) %>%
  filter(format(time, "%Y") == as.character(YEAR)) %>%
  group_by(geography) %>%
  summarize(cases = sum(fl_cyclo_cases, na.rm = TRUE) + sum(fl_cyclo_cases_probable, na.rm = TRUE), .groups = "drop")

oh <- vroom::vroom(file.path(cyclo_dir, "data/oh_cyclo/standard/data.csv.gz"), show_col_types = FALSE)
oh_county <- oh %>%
  mutate(geography = sprintf("%05d", geography)) %>%
  filter(format(time, "%Y") == as.character(YEAR)) %>%
  group_by(geography) %>%
  summarize(cases = sum(oh_cyclo_cases, na.rm = TRUE), .groups = "drop")

# Note: MI's cumulative count covers the Southeast Michigan outbreak period
# (since 2026-06-22 per MDHHS), not full-calendar-year Jan-Jun baseline cases.
mi <- vroom::vroom(file.path(cyclo_dir, "data/mi_cyclo/standard/data.csv.gz"), show_col_types = FALSE)
mi_county <- mi %>%
  mutate(geography = sprintf("%05d", geography)) %>%
  filter(format(time, "%Y") == as.character(YEAR), time == max(time)) %>%
  group_by(geography) %>%
  summarize(cases = sum(mi_cyclo_cases_cumulative, na.rm = TRUE), .groups = "drop")

county_cases <- bind_rows(fl_county, oh_county, mi_county)

# Latest reported date within YEAR for each state's county feed - these can
# differ from each other (and from `as_of`, which is just today's date).
state_asof <- c(
  FL = format(max(fl$time[format(fl$time, "%Y") == as.character(YEAR)]), "%b %d, %Y"),
  MI = format(max(mi$time[format(mi$time, "%Y") == as.character(YEAR)]), "%b %d, %Y"),
  OH = format(max(oh$time[format(oh$time, "%Y") == as.character(YEAR)]), "%b %d, %Y")
)

county_data <- pop_county %>%
  left_join(county_cases, by = "geography") %>%
  mutate(cases = ifelse(is.na(cases), 0, cases),
         cases_per_100k = cases / popsize * 100000)

# County shapefile for the FL/MI/OH panel map below (still a static ggplot
# map, so it needs sf geometry; the national map is an interactive plotly
# choropleth and only needs state abbreviations, not shapefiles).
options(tigris_use_cache = TRUE)
Sys.setenv(TIGRIS_CACHE_DIR = file.path(CACHE_DIR, "tigris_cache"))
dir.create(Sys.getenv("TIGRIS_CACHE_DIR"), showWarnings = FALSE, recursive = TRUE)

# WGS84 (not shift_geometry(), which reprojects to a non-geographic CRS for
# repositioning AK/HI/PR - unneeded here, and plotly needs plain lon/lat).
us_counties <- tigris::counties(state = names(state_names), cb = TRUE, resolution = "20m", year = 2022) %>%
  sf::st_transform(4326) %>%
  select(GEOID, NAMELSAD, STUSPS)

county_map <- us_counties %>% left_join(county_data, by = c("GEOID" = "geography"))
```

Cyclosporiasis is a diarrhea-causing infection from the parasite Cyclospira. It is spread through contaminated
food and water. While we typically see cases spike during the summer months, the CDC's National Notifiable Diseases Surveillance System (NNDSS) 
is reporting almost 2,000 cumulative cases this year, making it the largest outbreak in U.S. history, and cases continue to rise. 
In comparison, the U.S typically saw about half as many cases by this time in previous years. 

Cyclosporiasis is a nationally notifiable disease and is mandated to be 
reported in 47 states, but the exact total number of cases is difficult to pin down for several reasons. 
Local health departments are responsible for tracking and reporting case counts to states, which then report to the CDC. 
Additionally, not everyone with diarrhea seeks medical care, and diarrhea has many causes beyond cyclosporiasis, which complicates 
case tracking. This dashboard tracks reported cases, combining CDC's NNDSS data 
with additional data at the county level from three states: Florida,
Michigan, and Ohio. Data reflected below are current as of `r as_of`.

**Note**: State data are the authoritative source for case reporting. NNDSS data may lag or differ due to submission timing variations.

## Trends over time
National total shown by default; use the dropdown to switch to a specific state.

```{r}
#| label: trend-chart
#| fig-width: 10
#| fig-height: 9

nnds_national_weekly <- nnds %>%
  filter(geography == "00", mmwr_year >= 2022) %>%
  select(geography, mmwr_year, mmwr_week, cases = cyclosporiasis) %>%
  mutate(geography_name = "United States")

nnds_state_weekly <- nnds %>%
  filter(!is.na(geography), mmwr_year >= 2022) %>%
  select(geography, mmwr_year, mmwr_week, cases = cyclosporiasis) %>%
  inner_join(all_state_pop %>% select(geography, geography_name), by = "geography") %>%
  bind_rows(nnds_national_weekly) %>%
  # The source file is not guaranteed sorted by week within each
  # geography/year; plotly's "lines" mode draws points in row order (unlike
  # ggplot's geom_line, which auto-sorts by x), so an unsorted feed here
  # zig-zags into what looks like many lines instead of one clean line.
  arrange(geography_name, mmwr_year, mmwr_week)

states_sorted <- sort(unique(nnds_state_weekly$geography_name))
years_sorted  <- sort(unique(nnds_state_weekly$mmwr_year))
n_states <- length(states_sorted)
n_years  <- length(years_sorted)
default_state <- "United States"

# Kept to a single subtitle line, same reasoning as the national map's
# title: multi-line subtitles via <br> render with plotly's loose default
# line spacing ("double spaced"), and a longer text also risks overlapping
# the plot below without a very tall top margin. One short line avoids both.
make_trend_title <- function(st) list(
  text = paste0(
    "<b>Cumulative weekly cyclosporiasis cases</b><br>",
    "<span style='font-size:12px;color:#767676'>", st,
    ": cumulative NNDSS cases by calendar week, one line per year. Data from 2026 are preliminary.</span>"
  ),
  font = plotly_title_font, x = 0, xanchor = "left"
)

p_trend <- plot_ly()
for (st in states_sorted) {
  d_st <- nnds_state_weekly %>% filter(geography_name == st)
  for (i in seq_along(years_sorted)) {
    yr <- years_sorted[i]
    d_yr <- d_st %>% filter(mmwr_year == yr)
    is_default <- identical(st, default_state)
    p_trend <- p_trend %>% add_trace(
      data = d_yr, x = ~mmwr_week, y = ~cases, type = "scatter", mode = "lines+markers",
      name = as.character(yr), legendgroup = as.character(yr),
      showlegend = is_default, visible = is_default,
      line = list(color = pophive_line_colors[i],
                  dash = ifelse(pophive_line_types[i] == "dashed", "dash", "solid")),
      marker = list(color = pophive_line_colors[i], size = 5),
      hovertemplate = paste0(st, ", week %{x}: %{y} cases<extra>", yr, "</extra>")
    )
  }
}

# One dropdown button per state; toggles visible/showlegend for that state's
# block of `n_years` traces and hides every other state's traces.
state_buttons <- lapply(seq_along(states_sorted), function(si) {
  vis <- rep(FALSE, n_states * n_years)
  idx <- ((si - 1) * n_years + 1):(si * n_years)
  vis[idx] <- TRUE
  list(
    method = "update",
    args = list(list(visible = vis, showlegend = vis),
                list(title = make_trend_title(states_sorted[si]))),
    label = states_sorted[si]
  )
})

default_idx <- which(states_sorted == default_state) - 1

p_trend <- p_trend %>%
  layout(
    title = make_trend_title(default_state),
    # Dropdown and legend now sit INSIDE the plot's own top corners (same
    # coordinate space as the data, 0-1), not floating above it in the
    # margin - that "paper" region uses a different coordinate reference
    # than the title, which made them hard to align without overlapping.
    # Living inside the axes area instead guarantees no title collision.
    # Early calendar weeks are ~0 cases for every state/year, so the
    # top-left corner is always safe from real data; top-right similarly
    # only risks a thin sliver of a completed year's line.
    updatemenus = list(list(
      type = "dropdown", active = default_idx, x = 0.99, xanchor = "right", y = 0.99, yanchor = "top",
      font = plotly_body_font, buttons = state_buttons
    )),
    xaxis = list(title = "Calendar week", dtick = 4, tickfont = plotly_body_font,
                  titlefont = plotly_body_font, gridcolor = "white", zeroline = FALSE),
    yaxis = list(title = "Cases (cumulative YTD)", tickfont = plotly_body_font,
                  titlefont = plotly_body_font, gridcolor = "#e5e5e5", zeroline = FALSE,
                  griddash = "dash"),
    legend = list(orientation = "h", x = 0.01, y = 0.99, yanchor = "top", font = plotly_body_font),
    plot_bgcolor = "white", paper_bgcolor = "white",
    margin = list(t = 70, b = 70)
  ) %>%
  config(displayModeBar = FALSE)
p_trend
```

```{=html}
<details class="chart-info">
<summary>About this chart</summary>
<p>Map shows cumulative cyclosporiasis cases reported to NNDSS for selected year. 2026 data are preliminary and incomplete. Not all states mandate cyclosporiasis reporting; states with zero cases may lack reporting requirements rather than having no cases. Territories are excluded.
Standardized data can be found on our <a href="https://github.com/PopHIVE/Ingest/tree/main/data/nnds/standard">GitHub</a>. The raw NNDSS data from CDC are found <a href="https://www.cdc.gov/nndss/what-is-case-surveillance/">here</a>.</p>
</details>
```

## Geographic distribution

### Cases by state
2026 trends shown by default; use the slider to switch to an earlier year.

```{r}
#| label: national-map
#| fig-width: 9
#| fig-height: 6

annual_brks <- c(0, 0.1, 0.25, 0.5, 1, 2, 3, 5)

state_data_by_year <- state_data_by_year %>%
  mutate(hover_text = paste0(geography_name, ": ", scales::comma(cases),
                              " case", ifelse(cases == 1, "", "s"),
                              " (", sprintf("%.1f", cases_per_100k), " per 100,000)"))

z_max <- log1p(max(state_data_by_year$cases_per_100k, na.rm = TRUE))
default_year <- YEAR

# Title doesn't need the year - the year dropdown already makes that clear.
# Kept to a single subtitle line on purpose: multi-line subtitles (via <br>)
# render with plotly's fairly loose default line spacing, which reads as
# "double spaced" - one line avoids that entirely instead of fighting it.
map_title <- list(
  text = paste0(
    "<b>Cyclosporiasis Cases by State (NNDSS)</b><br>",
    "<span style='font-size:12px;color:#767676'>Cases per 100,000 population by state. 2026 are ",
    "through July 4. See county detail for FL, MI, OH below.</span>"
  ),
  font = plotly_title_font, x = 0, xanchor = "left"
)

p_national <- plot_ly()
for (yr in MAP_YEARS) {
  d <- state_data_by_year %>% filter(mmwr_year == yr)
  is_default <- yr == default_year
  p_national <- p_national %>% add_trace(
    type = "choropleth", locationmode = "USA-states",
    locations = d$state, z = log1p(d$cases_per_100k), zmin = 0, zmax = z_max,
    colorscale = pophive_magma_colorscale, text = d$hover_text, hoverinfo = "text",
    marker = list(line = list(color = "white", width = 0.5)), visible = is_default,
    # showscale = FALSE on the hidden year-traces matters: plotly reserves/
    # shrinks colorbar space accounting for EVERY trace that has a colorbar,
    # even invisible ones, so with 5 traces all requesting one, none of them
    # can ever reach the full len=1 height. Only the active trace claims it.
    showscale = is_default,
    colorbar = list(
      title = list(text = "Cases per<br>100,000", font = plotly_body_font),
      tickvals = log1p(annual_brks), ticktext = annual_brks,
      tickfont = plotly_body_font, len = 1, thickness = 18
    )
  )
}

# One slider step per year; toggles which single choropleth trace is
# visible, and moves showscale to that trace so the colorbar follows it.
# The title no longer depends on year, so steps don't need to touch it.
year_steps <- lapply(seq_along(MAP_YEARS), function(i) {
  vis <- rep(FALSE, length(MAP_YEARS))
  vis[i] <- TRUE
  list(method = "restyle", args = list(list(visible = vis, showscale = vis)), label = as.character(MAP_YEARS[i]))
})
default_year_idx <- which(MAP_YEARS == default_year) - 1

p_national <- p_national %>%
  layout(
    geo = list(scope = "usa", projection = list(type = "albers usa"), bgcolor = "transparent"),
    title = map_title,
    # Slider sits below the map (in the bottom margin) rather than a
    # dropdown floating over it.
    sliders = list(list(
      active = default_year_idx, pad = list(t = 30, l = 10),
      currentvalue = list(prefix = "Year: ", font = plotly_body_font, xanchor = "left"),
      font = plotly_body_font, steps = year_steps
    )),
    # Title is now just 2 short lines, so a smaller top margin both closes
    # the gap above it and - since colorbar len is a fraction of this plot
    # area - gives the len=1 colorbar more actual vertical room to fill.
    # margin.b is taller than the county maps' to leave the slider room.
    margin = list(t = 80, b = 90),
    paper_bgcolor = "white"
  ) %>%
  config(displayModeBar = FALSE)
p_national
```

```{=html}
<details class="chart-info">
<summary>About this chart</summary>
<p>Map shows cumulative cyclosporiasis cases reported to NNDSS by state for selected year. 2026 data are preliminary and incomplete. Not all states mandate cyclosporiasis reporting; states with zero cases may lack reporting requirements rather than having no cases. Territories are excluded.
Standardized data can be found on our <a href="https://github.com/PopHIVE/Ingest/tree/main/data/nnds/standard">GitHub</a>. The CDC presents its NNDSS data <a href="https://www.cdc.gov/nndss/what-is-case-surveillance/">here</a>.</p>
</details>
```

### FL, MI, and OH: county-level detail

These three states display case counts at the county level collected from state health departments. 
Hover over a county to see its case count. Most states only report cases at the state level. 
Oregon also reports at the county level, but its data are suppressed due to small sample sizes because Oregon is not currently experiencing a large outbreak.


```{r}
#| label: county-maps-setup
#| include: false

# plotly has no built-in US-counties location mode, so build a small GeoJSON
# from the tigris county shapes (FL/MI/OH only) and use it directly.
counties_geojson_file <- tempfile(fileext = ".geojson")
sf::st_write(county_map %>% select(GEOID), counties_geojson_file, driver = "GeoJSON", quiet = TRUE)
counties_geojson <- jsonlite::fromJSON(counties_geojson_file, simplifyVector = FALSE)

county_hover <- county_map %>%
  sf::st_drop_geometry() %>%
  mutate(hover_text = paste0(NAMELSAD, ": ", scales::comma(cases),
                              " case", ifelse(cases == 1, "", "s"),
                              " (", sprintf("%.1f", cases_per_100k), " per 100,000)"))

brks <- c(0, 1, 5, 25, 100)
z_max_county <- log1p(max(county_hover$cases_per_100k, na.rm = TRUE))

state_sources <- c(
  FL = paste0("Source: Florida Dept. of Health."),
  MI = paste0("Source: Michigan Dept. of Health & Human Services."),
  OH = paste0("Source: Ohio Dept. of Health.")
)

# FL's feed is explicitly preliminary (subject to revision for 1-2 weeks per
# its README); MI/OH are not flagged that way, hence the differing wording.
state_subtitles <- c(
  FL = paste0("Cumulative year-to-date preliminary cyclosporiasis  cases per 100,000 population as of ", state_asof[["FL"]], "."),
  MI = paste0("Cumulative preliminary cyclosporiasis  cases per 100,000 population reported during<br>the outbreak investigation (as of ", state_asof[["MI"]], ", starting July 1, 2026)."),
  OH = paste0("Cumulative year-to-date preliminary cyclosporiasis  cases per 100,000 population as of ", state_asof[["OH"]], ".")
)

make_county_plot <- function(abbr) {
  d <- county_hover %>% filter(STUSPS == abbr)
  plot_ly() %>%
    add_trace(
      type = "choropleth", geojson = counties_geojson, featureidkey = "properties.GEOID",
      locations = d$GEOID, z = log1p(d$cases_per_100k), zmin = 0, zmax = z_max_county,
      colorscale = pophive_magma_colorscale, text = d$hover_text, hoverinfo = "text",
      marker = list(line = list(color = "white", width = 0.4)),
      colorbar = list(
        title = list(text = "Cases per<br>100,000", font = plotly_body_font),
        tickvals = log1p(brks), ticktext = brks, tickfont = plotly_body_font,
        len = 0.8, thickness = 18
      )
    ) %>%
    layout(
      geo = list(scope = "usa", fitbounds = "locations", visible = FALSE),
      title = list(
        text = paste0(
          "<b>", state_names[[abbr]], "</b><br>",
          "<span style='font-size:12px;color:#767676'>", state_subtitles[[abbr]], "</span>"
        ),
        font = plotly_title_font, x = 0, xanchor = "left"
      ),
      annotations = list(list(
        text = state_sources[[abbr]], xref = "paper", yref = "paper",
        x = 0, y = -0.06, xanchor = "left", showarrow = FALSE, font = plotly_caption_font
      )),
      margin = list(t = 100, b = 40),
      paper_bgcolor = "white"
    ) %>%
    config(displayModeBar = FALSE)
}
```


```{r}
#| label: county-fl
#| fig-width: 9
#| fig-height: 7
make_county_plot("FL")
```


```{r}
#| label: county-mi
#| fig-width: 9
#| fig-height: 7
make_county_plot("MI")
```


```{r}
#| label: county-oh
#| fig-width: 9
#| fig-height: 7
make_county_plot("OH")
```

```{=html}
<details class="chart-info">
<summary>About these charts</summary>
<p>cyclosporiasis  cases are obtained from individual state health departments. Rates are calculated using county-specific population denominators from the 2021 census. Standardized data can be found on our <a href="https://github.com/PopHIVE/cyclo_resources/tree/main/data/oh_cyclo">GitHub</a>. Raw data from FL are found <a href="https://www.flhealthcharts.gov/ChartsReports/rdPage.aspx?rdReport=FrequencyMerlin.Frequency">here</a>; raw data from MI are found <a href="https://www.michigan.gov/mdhhs/keep-mi-healthy/infectious-diseases/infectious-disease-outbreaks">here</a>; and raw data from OH are found <a href="https://data.ohio.gov/wps/portal/gov/data/view/summary-of-infectious-diseases-in-ohio">here</a>.</p>
</details>
```

```{=html}
<footer class="pophive-footer">
  <p><strong>PopHIVE</strong> - Population Health Integrated Visualization Engine</p>
  <p>Yale School of Public Health &amp; Yale DISSC - Learn more at <a href="https://www.pophive.org/">pophive.org</a></p>
  <p class="disclaimer">Research tool for public health professionals. Not intended for clinical decision-making.</p>
</footer>
```
